### 2024-06-06 products/manuscript/Manuscript.qmd
inital commit for Part 1
+ LFS

---
title: "Crime in Austin Texas: Exploration, Classification, and Trends"
subtitle: ""
author: William Hyltin
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_ 
[Last year](https://www.fox7austin.com/news/austin-named-city-with-15th-biggest-homicide-rate-problem-study), Austin was ranked among the top US cities with a problem in homicide rate. Earlier [this year](https://www.kvue.com/article/news/local/austin-crime-rate-drops-2024-2020-comparison/269-b62042b7-a6f2-4224-87ff-ede655a37912), the claim was made that crime has dropped to a new low since 2021. The intention of this analysis will be to investigate both claims, and gain a better understanding of Austin's criminal activity.
What the analysis will not cover would be things like *causes* of the changes in crime, at least not directly. I may identify shifts that could inform the cause, but the goal will not be to drill down into every potential socioeconomic factor and pick the primary drivers, instead simply to observe what crime in the city of Austin looks like and what we might be able to expect going forward. Aside from resources outlined in our course resources, I was hoping to also draw from this [book](https://otexts.com/fpp2/index.html). It's not one I have used before but it had been recommended to me previously and I would like to learn more about time series and forecasting in general.

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
My primary data source will come from [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data), but I will likely need to pull from other sources as well, potentially finding other cities' crime stats that made the same list of problematic homicide rates, or even Austin's demographic information and statistics like population size.
```{r Load Raw Data}
d1 <- read.csv(here('data','raw-data','Crime_Reports_20240604.csv'))
```

```{r Quick Summaries}
summary(d1)
head(d1)
```
The data several data and time fields, many of which are character strings meaning they will need to be cleaned/recoded. The rest of the fields are mostly classification variables and location information, but with 2,470,867 there is quite a bit of opportunity for frequency and classification analysis.

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._
Ultimately, I would like to explore a few questions: 
  1. Has crime truly dropped, or is it expected to rise again as the year progresses? 
  2. Has the homicide rate dropped with crime? 
  3. Has the homicide rate been a problem for long time, or was it truly an emerging problem last year?

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References

### 2024-06-07 products/manuscript/Manuscript.qmd
minor updates to summary

@@
 The intention of this analysis will be to investigate both claims, and gain a better understanding of Austin's criminal activity.
+While these claims will be in mind throughout the steps of the analysis, I intend to approach it from more of an exploratory standpoint, so not to only go out and investigate the claims but really any insights that can be gleaned around Austin criminal activity and how it has changed over time.
 What the analysis will not cover would be things like *causes* of the changes in crime, at least not directly.
@@
 ```
-The data several data and time fields, many of which are character strings meaning they will need to be cleaned/recoded.
+The data has several data and time fields, many of which are character strings meaning they will need to be cleaned/recoded.
 The rest of the fields are mostly classification variables and location information, but with 2,470,867 there is quite a bit of opportunity for frequency and classification analysis.

### 2024-06-12 code/processing-code/processingfile.qmd
created processing file to start data cleaning and exploration. Added codebook for reference. renamed data file to be in line with naming conventions for Repo.

---
title: "An example cleaning script"
author: "Andreas Handel"
date: "2024-02-07"
output: html_document
editor: 
  markdown: 
    wrap: sentence
---

# Processing script

### Setup

Load needed packages.\
*Pacman::p_load* makes sure they are installed before auto-loading.\
*tidyverse* for general data processing/cleaning.\
*lubridate* for cleaning/re-coding date fields.\
*skimr* for nice visualization of data.\
*here* to set paths.

```{r}
#install.packages(pacman)
pacman::p_load(tidyverse,lubridate,skimr,here)
```

# Data loading

```{r}
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","raw_crime_060424.csv")
rawdata <- read.csv(data_location, TRUE)
```

# Check data

First we can look at the codebook, which can be found [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data).
I will recreate it here as it currently exists, and I will have a copy of the pdf attachment codebook included in the repo as well.
However, please note if this analysis is ever repeated that careful consideration should be taken to ensure field have not been changed, removed, or redefined.

| Field                       | Description                                                                                                  |
|------------------------------|------------------------------------------|
| Incident Number             | Incident report number                                                                                       |
| Highest Offense Description | Description                                                                                                  |
| Highest Offense Code        | Code                                                                                                         |
| Family Violence             | Incident involves family violence? Y = yes, N = no                                                           |
| Occurred Date Time          | Date and time (combined) incident occurred                                                                   |
| Occurred Date               | Date the incident occurred                                                                                   |
| Occurred Time               | Time the incident occurred                                                                                   |
| Report Date Time            | Date and time (combined) incident was reported                                                               |
| Report Date                 | Date the incident was reported                                                                               |
| Report Time                 | Time the incident was reported                                                                               |
| Location Type               | General description of the premise where the incident occurred                                               |
| Address                     | Incident location                                                                                            |
| Zip code                    | Zip code where incident occurred                                                                             |
| Council District            | Austin city council district where the incident occurred                                                     |
| APD Sector                  | APD sector where incident occurred                                                                           |
| APD District                | APD district where incident occurred                                                                         |
| PRA                         | APD police reporting area where incident occurred                                                            |
| Census Tract                | Census tract where incident occurred                                                                         |
| Clearance Status            | How/whether crime was solved (see Clearance lookup)                                                          |
| Clearance Date              | Date crime was solved                                                                                        |
| UCR Category                | Code for the most serious crimes identified by the FBI as part of its Uniform Crime Reporting program        |
| Category Description        | Description for the most serious crimes identified by the FBI as part of its Uniform Crime Reporting program |
| X-coordinate                | X-coordinate where the incident occurred                                                                     |
| Y-coordinate                | Y-coordinate where incident occurred                                                                         |
| Latitude                    | Latitude where incident occurred                                                                             |
| Longitude                   | Longitude where the incident occurred                                                                        |
| Location                    | 3rd party generated spatial column                                                                           |

| Clearance lookup |                      |
|------------------|----------------------|
| C                | Cleared by Arrest    |
| O                | Cleared by Exception |
| N                | Not cleared          |

Several ways of looking at the data

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
unique(rawdata$Location.Type)
```

# Cleaning

By inspecting the data as done above, we find some problems that need addressing:

```{r}
d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
                  dplyr::mutate(Height = as.numeric(Height))
skimr::skim(d1)
hist(d1$Height)
```

Now we see that there is one person with a height of 6.
That could be a typo, or someone mistakenly entered their height in feet.
Since we unfortunately don't know, we might need to remove this person, which we'll do here.

```{r}
d2 <- d1 %>% dplyr::mutate( Height = replace(Height, Height=="6",round(6*30.48,0)) )
skimr::skim(d2)
```

Height values seem ok now.

Now let's look at the `Weight` variable.
There is a person with weight of 7000, which is impossible, and one person with missing weight.
To be able to analyze the data, we'll remove those individuals as well.

```{r}
d3 <- d2 %>%  dplyr::filter(Weight != 7000) %>% tidyr::drop_na()
skimr::skim(d3)
```

Now checking the `Gender` variable.
Gender should be a categorical/factor variable but is loaded as character.
We can fix that with simple base R code to mix things up.

```{r}
d3$Gender <- as.factor(d3$Gender)  
skimr::skim(d3)
```

Now we see that there is another NA, but it's not `NA` from R, instead it was loaded as character and is now considered as a category.
Well proceed here by removing that individual with that NA entry.
Since this keeps an empty category for Gender, I'm also using droplevels() to get rid of it.

```{r}
d4 <- d3 %>% dplyr::filter( !(Gender %in% c("NA","N")) ) %>% droplevels()
skimr::skim(d4)
```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata <- d4
```

# Save data

Finally, we save the clean data as RDS file.
I suggest you save your processed and cleaned data as RDS or RDA/Rdata files.
This preserves coding like factors, characters, numeric, etc.
If you save as CSV, that information would get lost.
However, CSV is better for sharing with others since it's plain text.
If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data: http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

```{r}
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach.
It's often not the best.
based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

### 2024-06-13 code/processing-code/processingfile.qmd
Finished most of the date cleaning and documentation.

@@
 #install.packages(pacman)
-pacman::p_load(tidyverse,lubridate,skimr,here)
+pacman::p_load(tidyverse,lubridate,skimr,here, stringr)
 ```
@@
 # note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw-data","raw_crime_060424.csv")
+data_location <- here::here("data","raw-data","raw-crime-060424.csv")
 rawdata <- read.csv(data_location, TRUE)
@@
 
-By inspecting the data as done above, we find some problems that need addressing:
+By inspecting the data as done above, we find some problems that need addressing:\
+
+### Dates and Times
+
+We have six time related fields in this dataset: Occurred.Date.Time, Occurred.Date, Occurred.Time, Report.Date.Time, Report.Date, and Report.Time.\
+- The Date-Time and Date variables were read in as string, and the two Time variables were read in as numeric.
+So that we can treat these variables as actual representations of a point in time, we will need to re-code them into date, time, and date-time values.\
+- Many of the variables have missing or empty values where they shouldn't.
+For example, there are some instances where the Occurred.Date.Time variable is empty, but the Occurred.Date and Occurred.Time values are not.\
+- Finally, there are some instances where the time value does not make any sense.
+Values like 60 or 70 are impossible to occur and somewhat vague, so special consideration should be taken to account for these as well.
 
 ```{r}
-d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
-                  dplyr::mutate(Height = as.numeric(Height))
-skimr::skim(d1)
-hist(d1$Height)
+d1 <- rawdata %>% mutate(
+  test = (str_sub(as.character(Occurred.Time), -2, -1)
+          %>% as.numeric()),
+  Occurred.Date = mdy(Occurred.Date)
+  ,Report.Date = mdy(Report.Date)
+  ,Occurred.Date.Time = parse_date_time(Occurred.Date.Time,'%m/%d/%Y %I:%M:%S %p')
+  ,Report.Date.Time = parse_date_time(Report.Date.Time,'%m/%d/%Y %I:%M:%S %p')
+  ) %>% mutate(
+    Occurred.Time = if_else(
+      is.na(Occurred.Time) | test >= 60, 
+      if_else(
+        Occurred.Date == Report.Date,
+        if_else(
+          is.na(Report.Time), 
+          2359, Report.Time
+          ), 2359
+        ), if_else(
+          Occurred.Time == 2400, 
+          2359, Occurred.Time
+          ) 
+      ) %>% as.character()
+    ) %>% mutate(
+      Occurred.Time = paste0('000',Occurred.Time,'00') %>% 
+        str_sub(.,-6,-1) %>% 
+        parse_time('%H%M%S')
+      ) %>% select(-test)
 ```
@@
 based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).
+
+```{r Sandbox}
+#rawdata %>% filter(Occurred.Date.Time == '')
+
+d1 <- rawdata %>% mutate(
+  test = (str_sub(as.character(Occurred.Time), -2, -1)
+          %>% as.numeric()),
+  Occurred.Date = mdy(Occurred.Date)
+  ,Report.Date = mdy(Report.Date)
+  ,Occurred.Date.Time = parse_date_time(Occurred.Date.Time,'%m/%d/%Y %I:%M:%S %p')
+  ,Report.Date.Time = parse_date_time(Report.Date.Time,'%m/%d/%Y %I:%M:%S %p')
+  ) %>% mutate(
+    Occurred.Time = if_else(
+      is.na(Occurred.Time) | test >= 60, 
+      if_else(
+        Occurred.Date == Report.Date,
+        if_else(
+          is.na(Report.Time), 
+          2359, Report.Time
+          ), 2359
+        ), if_else(
+          Occurred.Time == 2400, 
+          2359, Occurred.Time
+          ) 
+      ) %>% as.character()
+    ) %>% mutate(
+      Occurred.Time = paste0('000',Occurred.Time,'00') %>% 
+        str_sub(.,-6,-1) %>% 
+        parse_time('%H%M%S')
+      ) %>% select(-test)
+```

### 2024-06-14 code/eda-code/eda.qmd
Significant amount more cleaning done. Created a processed data file and saved to data folder. Removed several template files.

version https://git-lfs.github.com/spec/v1
oid sha256:a063832bfc7ed6b54f7db32a75d484bdb6f59df9d1fba1d7ca6f90fe7911162f
size 4505

### 2024-06-14 code/processing-code/processingfile.qmd
Significant amount more cleaning done. Created a processed data file and saved to data folder. Removed several template files.

--- 2024-06-12 code/processing-code/processingfile.qmd
@@
----
-title: "An example cleaning script"
-author: "Andreas Handel"
-date: "2024-02-07"
-output: html_document
-editor: 
-  markdown: 
-    wrap: sentence
----
-
-# Processing script
-
-### Setup
-
-Load needed packages.\
-*Pacman::p_load* makes sure they are installed before auto-loading.\
-*tidyverse* for general data processing/cleaning.\
-*lubridate* for cleaning/re-coding date fields.\
-*skimr* for nice visualization of data.\
-*here* to set paths.
-
-```{r}
-#install.packages(pacman)
-pacman::p_load(tidyverse,lubridate,skimr,here)
-```
-
-# Data loading
-
-```{r}
-# path to data
-# note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw-data","raw_crime_060424.csv")
-rawdata <- read.csv(data_location, TRUE)
-```
-
-# Check data
-
-First we can look at the codebook, which can be found [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data).
-I will recreate it here as it currently exists, and I will have a copy of the pdf attachment codebook included in the repo as well.
-However, please note if this analysis is ever repeated that careful consideration should be taken to ensure field have not been changed, removed, or redefined.
-
-| Field                       | Description                                                                                                  |
-|------------------------------|------------------------------------------|
-| Incident Number             | Incident report number                                                                                       |
-| Highest Offense Description | Description                                                                                                  |
-| Highest Offense Code        | Code                                                                                                         |
-| Family Violence             | Incident involves family violence? Y = yes, N = no                                                           |
-| Occurred Date Time          | Date and time (combined) incident occurred                                                                   |
-| Occurred Date               | Date the incident occurred                                                                                   |
-| Occurred Time               | Time the incident occurred                                                                                   |
-| Report Date Time            | Date and time (combined) incident was reported                                                               |
-| Report Date                 | Date the incident was reported                                                                               |
-| Report Time                 | Time the incident was reported                                                                               |
-| Location Type               | General description of the premise where the incident occurred                                               |
-| Address                     | Incident location                                                                                            |
-| Zip code                    | Zip code where incident occurred                                                                             |
-| Council District            | Austin city council district where the incident occurred                                                     |
-| APD Sector                  | APD sector where incident occurred                                                                           |
-| APD District                | APD district where incident occurred                                                                         |
-| PRA                         | APD police reporting area where incident occurred                                                            |
-| Census Tract                | Census tract where incident occurred                                                                         |
-| Clearance Status            | How/whether crime was solved (see Clearance lookup)                                                          |
-| Clearance Date              | Date crime was solved                                                                                        |
-| UCR Category                | Code for the most serious crimes identified by the FBI as part of its Uniform Crime Reporting program        |
-| Category Description        | Description for the most serious crimes identified by the FBI as part of its Uniform Crime Reporting program |
-| X-coordinate                | X-coordinate where the incident occurred                                                                     |
-| Y-coordinate                | Y-coordinate where incident occurred                                                                         |
-| Latitude                    | Latitude where incident occurred                                                                             |
-| Longitude                   | Longitude where the incident occurred                                                                        |
-| Location                    | 3rd party generated spatial column                                                                           |
-
-| Clearance lookup |                      |
-|------------------|----------------------|
-| C                | Cleared by Arrest    |
-| O                | Cleared by Exception |
-| N                | Not cleared          |
-
-Several ways of looking at the data
-
-```{r}
-dplyr::glimpse(rawdata)
-summary(rawdata)
-head(rawdata)
-skimr::skim(rawdata)
-unique(rawdata$Location.Type)
-```
-
-# Cleaning
-
-By inspecting the data as done above, we find some problems that need addressing:
-
-```{r}
-d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
-                  dplyr::mutate(Height = as.numeric(Height))
-skimr::skim(d1)
-hist(d1$Height)
-```
-
-Now we see that there is one person with a height of 6.
-That could be a typo, or someone mistakenly entered their height in feet.
-Since we unfortunately don't know, we might need to remove this person, which we'll do here.
-
-```{r}
-d2 <- d1 %>% dplyr::mutate( Height = replace(Height, Height=="6",round(6*30.48,0)) )
-skimr::skim(d2)
-```
-
-Height values seem ok now.
-
-Now let's look at the `Weight` variable.
-There is a person with weight of 7000, which is impossible, and one person with missing weight.
-To be able to analyze the data, we'll remove those individuals as well.
-
-```{r}
-d3 <- d2 %>%  dplyr::filter(Weight != 7000) %>% tidyr::drop_na()
-skimr::skim(d3)
-```
-
-Now checking the `Gender` variable.
-Gender should be a categorical/factor variable but is loaded as character.
-We can fix that with simple base R code to mix things up.
-
-```{r}
-d3$Gender <- as.factor(d3$Gender)  
-skimr::skim(d3)
-```
-
-Now we see that there is another NA, but it's not `NA` from R, instead it was loaded as character and is now considered as a category.
-Well proceed here by removing that individual with that NA entry.
-Since this keeps an empty category for Gender, I'm also using droplevels() to get rid of it.
-
-```{r}
-d4 <- d3 %>% dplyr::filter( !(Gender %in% c("NA","N")) ) %>% droplevels()
-skimr::skim(d4)
-```
-
-All done, data is clean now.
-
-Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.
-
-```{r}
-processeddata <- d4
-```
-
-# Save data
-
-Finally, we save the clean data as RDS file.
-I suggest you save your processed and cleaned data as RDS or RDA/Rdata files.
-This preserves coding like factors, characters, numeric, etc.
-If you save as CSV, that information would get lost.
-However, CSV is better for sharing with others since it's plain text.
-If you do CSV, you might want to write down somewhere what each variable is.
-
-See here for some suggestions on how to store your processed data: http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata
-
-```{r}
-save_data_location <- here::here("data","processed-data","processeddata.rds")
-saveRDS(processeddata, file = save_data_location)
-```
-
-# Notes
-
-Removing anyone who had "faulty" or missing data is one approach.
-It's often not the best.
-based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).
+version https://git-lfs.github.com/spec/v1
+oid sha256:8ffd2c5fe0b3ccdc979dec853978e42ac882fefa8612df567f1f65d7abde7670
+size 14120

### 2024-06-14 products/manuscript/Manuscript.qmd
created processing file to start data cleaning and exploration. Added codebook for reference. renamed data file to be in line with naming conventions for Repo.
Significant amount more cleaning done. Created a processed data file and saved to data folder. Removed several template files.

--- 2024-06-06 products/manuscript/Manuscript.qmd
@@
----
-title: "Crime in Austin Texas: Exploration, Classification, and Trends"
-subtitle: ""
-author: William Hyltin
-date: "`r Sys.Date()`"
-format:
-  docx:
-    toc: false
-    number-sections: true
-    highlight-style: github
-bibliography: ../../assets/dataanalysis-references.bib
-csl: ../../assets/apa.csl
----
-
-The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
-```{r, echo=FALSE, message=FALSE}
-# load a few R packages
-library(here)
-library(knitr)
-```
-
-# Summary/Abstract
-_Write a summary of your project._
-
-{{< pagebreak >}}
-
-# Introduction
-
-## General Background Information
-_Provide enough background on your topic that others can understand the why and how of your analysis_ 
-[Last year](https://www.fox7austin.com/news/austin-named-city-with-15th-biggest-homicide-rate-problem-study), Austin was ranked among the top US cities with a problem in homicide rate.
-Earlier [this year](https://www.kvue.com/article/news/local/austin-crime-rate-drops-2024-2020-comparison/269-b62042b7-a6f2-4224-87ff-ede655a37912), the claim was made that crime has dropped to a new low since 2021.
-The intention of this analysis will be to investigate both claims, and gain a better understanding of Austin's criminal activity.
-What the analysis will not cover would be things like *causes* of the changes in crime, at least not directly.
-I may identify shifts that could inform the cause, but the goal will not be to drill down into every potential socioeconomic factor and pick the primary drivers, instead simply to observe what crime in the city of Austin looks like and what we might be able to expect going forward.
-Aside from resources outlined in our course resources, I was hoping to also draw from this [book](https://otexts.com/fpp2/index.html).
-It's not one I have used before but it had been recommended to me previously and I would like to learn more about time series and forecasting in general.
-
-## Description of data and data source
-_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
-My primary data source will come from [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data), but I will likely need to pull from other sources as well, potentially finding other cities' crime stats that made the same list of problematic homicide rates, or even Austin's demographic information and statistics like population size.
-```{r Load Raw Data}
-d1 <- read.csv(here('data','raw-data','Crime_Reports_20240604.csv'))
-```
-
-```{r Quick Summaries}
-summary(d1)
-head(d1)
-```
-The data several data and time fields, many of which are character strings meaning they will need to be cleaned/recoded.
-The rest of the fields are mostly classification variables and location information, but with 2,470,867 there is quite a bit of opportunity for frequency and classification analysis.
-
-## Questions/Hypotheses to be addressed
-_State the research questions you plan to answer with this analysis._
-Ultimately, I would like to explore a few questions: 
-  1. Has crime truly dropped, or is it expected to rise again as the year progresses? 
-  2. Has the homicide rate dropped with crime? 
-  3. Has the homicide rate been a problem for long time, or was it truly an emerging problem last year?
-
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key.
-Then you can include like this:
-
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].
-
-{{< pagebreak >}}
-
-# Methods
-
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
-
-## Schematic of workflow
-
-Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-@fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
-We store those figures in the `assets` folder.
-
-```{r}
-#| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
-#| echo: FALSE
-knitr::include_graphics(here("assets","antigen-recognition.png"))
-```
-
-## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
-
-## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
-
-## Statistical analysis
-_Explain anything related to your statistical analyses._
-
-{{< pagebreak >}}
-
-# Results
-
-## Exploratory/Descriptive analysis
-
-_Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
-
-@tbl-summarytable shows a summary of the data.
-
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-I generally recommend the `here` package.
-
-```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
-#| echo: FALSE
-resulttable=readRDS("../../results/tables/summarytable.rds")
-knitr::kable(resulttable)
-```
-
-## Basic statistical analysis
-
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
-#| echo: FALSE
-knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
-```
-
-## Full analysis
-
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
-knitr::kable(resulttable2)
-```
-
-{{< pagebreak >}}
-
-# Discussion
-
-## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
-
-## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
-
-## Conclusions
-_What are the main take-home messages?_
-
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like.
-
-{{< pagebreak >}}
-
-# References
+version https://git-lfs.github.com/spec/v1
+oid sha256:53791af84dddf4d0e7fd1fcad04d4f30f60de0ca33f13bf28575f0f87d12c65f
+size 4288

### 2024-06-14 code/analysis-code/statistical-analysis.R

@@ -1,56 +1,3 @@
-###############################
-# analysis script
-#
-#this script loads the processed, cleaned data, does a simple analysis
-#and saves the results to the results folder
-
-#load needed packages. make sure they are installed.
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
-
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed-data","processeddata.rds")
-
-#load data. 
-mydata <- readRDS(data_location)
-
-
-######################################
-#Data fitting/statistical analysis
-######################################
-
-############################
-#### First model fit
-# fit linear model using height as outcome, weight as predictor
-
-lmfit1 <- lm(Height ~ Weight, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable1 <- broom::tidy(lmfit1)
-
-#look at fit results
-print(lmtable1)
-
-# save fit results table  
-table_file1 = here("results", "tables", "resulttable1.rds")
-saveRDS(lmtable1, file = table_file1)
-
-############################
-#### Second model fit
-# fit linear model using height as outcome, weight and gender as predictor
-
-lmfit2 <- lm(Height ~ Weight + Gender, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable2 <- broom::tidy(lmfit2)
-
-#look at fit results
-print(lmtable2)
-
-# save fit results table  
-table_file2 = here("results", "tables", "resulttable2.rds")
-saveRDS(lmtable2, file = table_file2)
-
-  
+version https://git-lfs.github.com/spec/v1
+oid sha256:2e185b3e8e00e0617e772d144729b4161ff6784d1f63d08966bb2d951bf473ac
+size 1512

### 2024-06-14 code/eda-code/edacode.R

@@ -1,44 +1,3 @@
-## ---- packages --------
-#load needed packages. make sure they are installed.
-library(here) #for data loading/saving
-library(dplyr)
-library(skimr)
-library(ggplot2)
-
-## ---- loaddata --------
-#Path to data. Note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed-data","processeddata.rds")
-#load data
-mydata <- readRDS(data_location)
-
-## ---- table1 --------
-summary_df = skimr::skim(mydata)
-print(summary_df)
-# save to file
-summarytable_file = here("results","tables", "summarytable.rds")
-saveRDS(summary_df, file = summarytable_file)
-
-## ---- height --------
-p1 <- mydata %>% ggplot(aes(x=Height)) + geom_histogram() 
-plot(p1)
-figure_file = here("results", "figures", "height-distribution.png")
-ggsave(filename = figure_file, plot=p1) 
-
-## ---- weight --------
-p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
-plot(p2)
-figure_file = here("results", "figures", "weight-distribution.png")
-ggsave(filename = figure_file, plot=p2) 
-
-## ---- fitfig1 --------
-p3 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
-plot(p3)
-figure_file = here("results","figures", "height-weight.png")
-
-## ---- fitfig2 --------
-p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Gender)) + geom_point() + geom_smooth(method='lm')
-plot(p4)
-figure_file = here("results","figures", "height-weight-stratified.png")
-ggsave(filename = figure_file, plot=p4) 
-
-
+version https://git-lfs.github.com/spec/v1
+oid sha256:0b112866218719c0d0f5936d491afb5c64f891b611f3d518c5d353b422a0e86f
+size 1432

### 2024-06-14 products/manuscript/Manuscript.qmd
repointed file directory in manusripts, added some notes

@@
-version https://git-lfs.github.com/spec/v1
-oid sha256:53791af84dddf4d0e7fd1fcad04d4f30f60de0ca33f13bf28575f0f87d12c65f
-size 4288
+---
+title: "Crime in Austin Texas: Exploration, Classification, and Trends"
+subtitle: ""
+author: William Hyltin
+date: "`r Sys.Date()`"
+format:
+  docx:
+    toc: false
+    number-sections: true
+    highlight-style: github
+bibliography: ../../assets/dataanalysis-references.bib
+csl: ../../assets/apa.csl
+---
+
+```{r Setup, echo=FALSE, message=FALSE}
+# load a few R packages
+library(here)
+library(knitr)
+```
+
+# Summary/Abstract
+_Write a summary of your project._
+
+{{< pagebreak >}}
+
+# Introduction
+
+## General Background Information
+[Last year](https://www.fox7austin.com/news/austin-named-city-with-15th-biggest-homicide-rate-problem-study), Austin was ranked among the top US cities with a problem in homicide rate.
+Earlier [this year](https://www.kvue.com/article/news/local/austin-crime-rate-drops-2024-2020-comparison/269-b62042b7-a6f2-4224-87ff-ede655a37912), the claim was made that crime has dropped to a new low since 2021.
+The intention of this analysis will be to investigate both claims, and gain a better understanding of Austin's criminal activity.
+While these claims will be in mind throughout the steps of the analysis, I intend to approach it from more of an exploratory standpoint, so not to only go out and investigate the claims but really any insights that can be gleaned around Austin criminal activity and how it has changed over time.
+What the analysis will not cover would be things like *causes* of the changes in crime, at least not directly.
+I may identify shifts that could inform the cause, but the goal will not be to drill down into every potential socioeconomic factor and pick the primary drivers, instead simply to observe what crime in the city of Austin looks like and what we might be able to expect going forward.
+Aside from resources outlined in our course resources, I was hoping to also draw from this [book](https://otexts.com/fpp2/index.html).
+It's not one I have used before but it had been recommended to me previously and I would like to learn more about time series and forecasting in general.
+
+## Description of data and data source
+My primary data source will come from [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data), but I will likely need to pull from other sources as well, potentially finding other cities' crime stats that made the same list of problematic homicide rates, or even Austin's demographic information and statistics like population size.
+
+The raw data has several data and time fields, many of which are character strings meaning they will need to be cleaned/recoded.
+The rest of the fields are mostly classification variables and location information with some minor cleaning opportunities, but with 2,470,867 there is quite a bit of opportunity for frequency and classification analysis.
+After cleaning the data (done in a separate file found in this repo) we can take a look a more meaningful look.
+
+```{r Load Processed Data}
+d1 <- readRDS(here('data','processed-data','processedcrime.rds'))
+```
+
+```{r Quick Summaries}
+summary(d1)
+head(d1)
+```
+
+## Questions/Hypotheses to be addressed
+Ultimately, I would like to explore a few questions: 
+  1. Has crime truly dropped, or is it expected to rise again as the year progresses? 
+  2. Has the homicide rate dropped with crime? 
+  3. Has the homicide rate been a problem for long time, or was it truly an emerging problem last year?
+
+{{< pagebreak >}}
+
+# Methods
+
+_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
+
+## Schematic of workflow
+
+```{r}
+
+```
+
+## Data aquisition
+
+## Statistical analysis
+
+{{< pagebreak >}}
+
+# Results
+
+## Exploratory/Descriptive analysis
+
+```{r}
+
+```
+
+## Basic statistical analysis
+
+@fig-result shows a scatterplot figure produced by one of the R scripts.
+
+```{r}
+```
+
+## Full analysis
+
+```{r}
+
+```
+
+{{< pagebreak >}}
+
+# Discussion
+
+## Summary and Interpretation
+
+## Strengths and Limitations
+
+## Conclusions
+
+{{< pagebreak >}}
+
+# References

### 2024-06-15 products/manuscript/Manuscript.qmd
Updated some of the Readmes, explored data with some visualizations in Manuscript

@@
 # load a few R packages
-library(here)
-library(knitr)
+pacman::p_load(here, knitr, tidyverse)
 ```
@@
   3. Has the homicide rate been a problem for long time, or was it truly an emerging problem last year?
+
+```{r}
+d1 %>% group_by(Occurred.Date) %>% 
+  summarize(
+  cnt = n()
+) %>% as.data.frame() %>% 
+  ggplot(mapping = aes(x=Occurred.Date, y=cnt)) +
+  geom_line()
+```
+```{r}
+d1 %>% 
+  mutate(years = year(Occurred.Date),
+         years = as.character(years)
+         ) %>% 
+  filter(years == '2023' | years == '2022') %>% 
+  group_by(Family.Violence, Zip.Code, years) %>% 
+  summarize(cnt = n(), .groups = 'drop') %>% 
+  group_by(Zip.Code, years) %>% 
+  mutate(tot = sum(cnt)) %>% 
+  relocate(tot, .after = years) %>% 
+  mutate(Zip.Code = as.character(Zip.Code)) %>%
+  as.data.frame() %>% 
+  arrange(desc(tot)) %>% head(100) %>% 
+  ggplot(aes(fill=Family.Violence, x= reorder(Zip.Code,cnt), y=cnt)) +
+  geom_bar(position = 'stack', stat='identity') +
+  facet_wrap(~years) +
+  coord_flip()
+```
 

### 2024-06-22 code/eda-code/eda.qmd
Lots of exploratory work done, deep dive into time series data, and many of the categorical variables in eda code. Moved a lot of that work out of the manuscript. Added descriptions, ideas, and interpretations. Started working on an ARIMA model for crime frequency.

@@
-version https://git-lfs.github.com/spec/v1
-oid sha256:a063832bfc7ed6b54f7db32a75d484bdb6f59df9d1fba1d7ca6f90fe7911162f
-size 4505
+---
+title: "An example exploratory analysis script"
+date: "2024-02-07"
+output: html_document
+---
+
+# Setup
+
+```{r setup}
+#load needed packages. make sure they are installed.
+pacman::p_load(here, knitr, tidyverse, skimr, fpp2)
+theme_set(theme_minimal())
+```
+
+Load the data.
+
+```{r}
+d1 <- readRDS(here('data','processed-data','processed-crime.rds'))
+zips <- read_csv(here('data','raw-data','austin-zip-codes.csv'))
+```
+
+# Data exploration through tables
+
+Showing a bit of code to produce and save a summary table.
+
+```{r}
+summary(d1)
+head(d1)
+str(d1)
+skim(d1)
+```
+
+```{r}
+summary(zips)
+head(zips)
+str(zips)
+skim(zips)
+```
+
+Important to note that the dataset has several date fields, and that alone will be ripe with opportunity for exploration.
+However I want to take a look at some of the other variables first, to see if maybe isolating to one of the other categories, or including other categorical variables in the time series could be useful.
+First off is Zip Code, because location is often a major factor in determining the amount and type of crime that would occur.
+
+```{r}
+d1 %>% 
+  mutate(years = year(Occurred.Date),
+         years = as.character(years)
+         ) %>% 
+  group_by(years, Zip.Code) %>% 
+  summarize(
+    count = n()
+  ) %>% as.data.frame() %>% pivot_wider(names_from = years, values_from = count)
+```
+
+Layering in the year gives an interesting view, because it gives us the opportunity to see major increases in certain locations over time.
+However, there are a lot of zip Codes and a lot of years in this dataset.
+A visualization may help with this, my first thought being a chloropeth plot with a slider for the year value.
+Other quicker solutions might be looking at the individual years as observations of sorts, then looking at summary statistics.
+More to come here.
+
+```{r}
+d1 %>% 
+  mutate(years = year(Occurred.Date)
+         #years = as.character(years)
+         ) %>% 
+  #filter(years == '2023' | years == '2022') %>% 
+  #group_by(Family.Violence, Highest.Offense.Description, years) %>% 
+  group_by(Highest.Offense.Description, years) %>% 
+  summarize(cnt = n(), .groups = 'drop') %>% 
+  #group_by(Highest.Offense.Description, years) %>% 
+  #mutate(tot = sum(cnt)) %>% 
+  #relocate(tot, .after = years) %>% 
+  mutate(Highest.Offense.Description = as.factor(Highest.Offense.Description)) %>%
+  as.data.frame() %>%
+  arrange(years) %>% 
+  pivot_wider(names_from = years, values_from = cnt)
+
+```
+
+Similar to our zip code variable here, but there are far too many crime descriptions to use.
+Grouping some together may work, I don't think we have enough information in the dataset overall to do any sort of classification modeling however, so it may amount to a rules based classification using regular expressions.
+Other variables may still help with this, such as the UCF classification variables.
+
+```{r}
+d1 %>% 
+  mutate(
+    Category.Description = if_else(
+      Category.Description == '', 'NA/Unknown', Category.Description
+      )
+    ) %>% 
+  group_by(Category.Description, Highest.Offense.Description) %>% 
+  summarize(
+    cnt = n()
+    ) %>% 
+  pivot_wider(
+    names_from = Category.Description, values_from = cnt
+    )
+```
+
+The UCF classifiers are much more consumable, but there are more incidents that are not classified than are.
+That said, this might be a good start for a rules based classification.
+Many of the unclassified incidents would not fit very nicely into the groups anyway, so some new categories may need to be made.
+I may or may not come back to the re-classification part, but ultimately the genesis of this project was with reports on the amount of homicide in the city, so the already existing descriptions may be enough to drill down into just those.
+
+*reminder to myself to save figures from this file to bring into manuscript and presentation later.*
+
+# Data exploration through figures
+
+Time is a variable that is much easier to interpret with a plot, so let's start there.
+Time series first.
+
+```{r}
+d1 %>% group_by(Occurred.Date) %>% 
+  summarize(
+  cnt = n()
+) %>% as.data.frame() %>% 
+  ggplot(mapping = aes(x=Occurred.Date, y=cnt)) +
+  geom_line()
+```
+
+I expected aggregating by day to actually look worse than this.
+There is a clear trend, upwards at first but then starts decreasing around 2008 or so.
+Seeing this I am tempted to looking into time series modeling.
+There may be a seasonal pattern in some of those spikes, but that may be something that can be resolved with different methods.
+I'm tempted to go ARIMA, because frankly that's the one I'm familiar with, but I'll have to do some reading up on my options.
+An interesting idea might be to fit the model, make predictions, and then see where I landed compared to the actual data before the class ends.
+
+```{r}
+d1 %>% filter(Occurred.Date <= '2023-12-31') %>%
+  mutate(Occurred.Date = trunc.Date(Occurred.Date, 'months')) %>% 
+  group_by(Occurred.Date) %>% 
+  summarize(
+  cnt = n()
+) %>% as.data.frame() %>% 
+  ggplot(mapping = aes(x=Occurred.Date, y=cnt)) +
+  geom_line()
+```
+
+Monthly looks even better, again further lending itself to the idea of a time series model.
+The downward spikes are more prevalent here and may cause problems themselves as well, but again the seasonality may be able to be resolved with things like logs or differencing.
+As a note, in this and the last plot I filter out 2024, mostly due to a large drop at the end from the incomplete month, but also for the aforementioned idea of comparing predictions to actual results.
+
+A potentially quick way to identify "peak" months for crime would be via a bar chart aggregating all occurrences in each month together.
+I was hoping to see some quick trends here, but I see surprisingly little here.
+The highest month is May, potentially because of school letting out/ graduations for UT Austin, but I wouldnt say the difference is large enough compared to other months to really point to any one reason.
+This may be another one to layer in with time, like a chart with a slider for the year to see shifts in crime frequency by month over time.
+
+```{r}
+d2 <- d1 %>% 
+  mutate(
+    Occurred.Month = month(Occurred.Date, label = TRUE),
+    Occurred.Day = weekdays(Occurred.Date),
+    Occurred.Year = year(Occurred.Date),
+    Occur.Report.Diff = Report.Date - Occurred.Date
+  )
+d2 %>% 
+  #filter(Occurred.Year == 2023) %>%
+  ggplot(aes(x=Occurred.Month)) +
+  geom_bar()
+```
+
+Now a Zip Code bar chart.
+Really a bit of a graphical representation of what we saw earlier, though I filtered it to the two most recent complete years.
+Again, looking for big swings, and simultaneously taking a quick peak at the Family Violence indicator.
+Surprisingly low volume there, which leads me to believe it just doesn't get captured appropriately every time, especially when comparing with the descriptions themselves, which we will see in a moment.
+Going back to the zip codes, there might be an opportunity to do something like binning the zip codes with high medium and low crime areas, or something of the sort.
+
+```{r}
+d1 %>% 
+  mutate(years = year(Occurred.Date),
+         years = as.character(years)
+         ) %>% 
+  filter(years == '2023' | years == '2022') %>% 
+  group_by(Family.Violence, Zip.Code, years) %>% 
+  summarize(cnt = n(), .groups = 'drop') %>% 
+  group_by(Zip.Code, years) %>% 
+  mutate(tot = sum(cnt)) %>% 
+  relocate(tot, .after = years) %>% 
+  mutate(Zip.Code = as.character(Zip.Code)) %>%
+  as.data.frame() %>% 
+  arrange(desc(tot)) %>% head(100) %>% 
+  ggplot(aes(fill=Family.Violence, x= reorder(Zip.Code,cnt), y=cnt)) +
+  geom_bar(position = 'stack', stat='identity') +
+  facet_wrap(~years) +
+  coord_flip()
+```
+
+Now a bar chart by crime description.
+This is similar to the last plot, but a little more disparate here.
+There are so many more crime descriptions than zip codes, so they are much more easily spread thin.
+That said, that actually makes the top five crime descriptions stand out a little more.
+So going back to the idea of creating new classifications, that may be a good idea to make sure those 5 are appropriately captured.
+And again, the family violence indicator is captured almost entirely by one description, yet the top description is Family disturbance.
+An easy explanation may be because the description is of the Highest Offense in the incident, so when there is physical family violence, it is often the highest offense.
+
+```{r}
+d1 %>% 
+  mutate(years = year(Occurred.Date),
+         years = as.character(years)
+         ) %>% 
+  filter(years == '2023' | years == '2022') %>% 
+  group_by(Family.Violence, Highest.Offense.Description, years) %>% 
+  summarize(cnt = n(), .groups = 'drop') %>% 
+  group_by(Highest.Offense.Description, years) %>% 
+  mutate(tot = sum(cnt)) %>% 
+  relocate(tot, .after = years) %>% 
+  mutate(Highest.Offense.Description = as.factor(Highest.Offense.Description)) %>%
+  as.data.frame() %>% 
+  arrange(desc(tot)) %>% head(50) %>% 
+  ggplot(aes(fill=Family.Violence, x= reorder(Highest.Offense.Description,cnt), y=cnt)) +
+  geom_bar(position = 'stack', stat='identity') +
+  facet_wrap(~years) +
+  coord_flip() 
+```
+
+# Notes
+
+Including this sandbox chunk where I can experiment with code as needed.
+ARIMA is not going well, mostly due to seasonal nature of data. Filtering the time range might help, or isolating other variables to do things liek predict certain types of crime instead.
+
+```{r sandbox}
+
+d1 %>% mutate(Category.Description = if_else(Category.Description == '', 'NA/Unknown', Category.Description)) %>% group_by(Category.Description, Highest.Offense.Description) %>% summarize(cnt = n()) %>% pivot_wider(names_from = Category.Description, values_from = cnt)
+
+temp <- d1 %>%
+         mutate(Occurred.Date = trunc.Date(Occurred.Date, 'months')) %>% 
+         group_by(Occurred.Date) %>% 
+         summarize(
+             cnt = n()
+         ) %>% as.data.frame() %>% 
+  filter(year(Occurred.Date) != 2024)
+rownames(temp) <- temp$Occurred.Date
+temp <- temp %>% select(cnt)
+
+temp2 <- temp %>% as.ts()
+Acf(temp2)
+Box.test(temp2, lag = 12, type = 'Ljung-Box')
+auto.arima(temp2)
+temp2 %>% log() %>% diff(12) %>% diff() %>% diff() %>% diff() %>% diff() %>% diff() %>% Acf()
+```

### 2024-06-22 code/processing-code/processingfile.qmd
Lots of exploratory work done, deep dive into time series data, and many of the categorical variables in eda code. Moved a lot of that work out of the manuscript. Added descriptions, ideas, and interpretations. Started working on an ARIMA model for crime frequency.

--- 2024-06-13 code/processing-code/processingfile.qmd
@@
 ---
-title: "An example cleaning script"
-author: "Andreas Handel"
-date: "2024-02-07"
+title: "Cleaning up Crime"
+author: "William Hyltin"
+date: today
 output: html_document
@@
 
-We have six time related fields in this dataset: Occurred.Date.Time, Occurred.Date, Occurred.Time, Report.Date.Time, Report.Date, and Report.Time.\
-- The Date-Time and Date variables were read in as string, and the two Time variables were read in as numeric.
+We have seven time related fields in this dataset: Occurred.Date.Time, Occurred.Date, Occurred.Time, Report.Date.Time, Report.Date, Report.Time, and Clearance.Date.\
+- The Date-Time and Date variables were read in as strings, and the two Time variables were read in as numeric.
 So that we can treat these variables as actual representations of a point in time, we will need to re-code them into date, time, and date-time values.\
@@
   ,Report.Date.Time = parse_date_time(Report.Date.Time,'%m/%d/%Y %I:%M:%S %p')
+  ,Clearance.Date = mdy(Clearance.Date)
+  ) %>% mutate(
+    Occurred.Time = if_else(
+      is.na(Occurred.Time) | test >= 60, 
+      if_else(
+        Occurred.Date == Report.Date,
+        if_else(
+          is.na(Report.Time), 
+          2359, Report.Time
+          ), 2359
+        ), if_else(
+          Occurred.Time == 2400, 
+          2359, Occurred.Time
+          ) 
+      ) %>% as.character()
+    ) %>% mutate(
+      Occurred.Time = paste0('000',Occurred.Time,'00') %>% 
+        str_sub(.,-6,-1) %>% 
+        parse_time('%H%M%S')
+      ,Occurred.Date.Time = if_else(
+        is.na(Occurred.Date.Time), 
+        ymd_hms(paste(Occurred.Date, Occurred.Time)), Occurred.Date.Time
+        )
+      ,Report.Time = if_else(
+        is.na(Report.Time), 2359, Report.Time) %>%
+        as.character() %>% 
+        paste0('000',.,'00') %>% 
+        str_sub(.,-6,-1) %>% 
+        parse_time('%H%M%S')
+      ,Report.Date.Time = if_else(
+        is.na(Report.Date.Time),
+        ymd_hms(paste(Report.Date, Report.Time)), Report.Date.Time
+      )
+      ) %>% select(-test)
+skim(d1)
+```
+
+This should address all of the time-based variables.
+I may come back and update some of this, solely because I made some more analytic decisions regarding the empty values for Occurred.Date.Time.
+The logic is that if the time is missing, the occurrence would have had to have happened at least before the report time, though if the report occurred on a different day we assume the occurrence would have to be at least by midnight the same day.
+The number of empties, once the three "Occurred" variables are coalesced where appropriate, are relatively low, so I don't expect this to have significant effect on any results either way.
+
+### Categorical Variables
+
+In the initial looks at the dataset a few columns stand out as potential factor variables.  
+ - Family Violence is a mostly complete field, though a minor oddity to it is the existence of a third factor when the codebook says it should only be two.  
+ - APD Sector, APD District, and PRA may be the best values to serve as a categorical representation of location, since it has relatively few missing variables.
+Zip code may also work and be more consumable to the average person, however it does have more missing than the others mentioned.
+Lastly, X and Y Coordinates, though not naturally categorical, may be able to be used to create categorical groupings considering their completeness.
+ - Location Type is another good categorical variable.
+There are a lot of categories, but with one like this it may be easy to combine some of them.
+Empty values are climbing, but still less than 1% of the data.
+ - Highest Offense code has 396 distinct values, which would be far too many for categorical use, but again may be able to be grouped.
+ - Clearance Status sounds like a very interesting variable to use, however it is one of the variables with the highest number of empties/missing.
+Furthermore, it's confounded by Clearance Date, which has fewer (but still a lot of) missing and empty values.
+This is one that may take some outside research, on what constitutes a "clearance" and what it means in relation to a clearance date.
+ - UCR Category and Description also sound like very interesting variables, but are the worst offenders in terms of missing data.
+This makes sense from info from the codebook, that it is reserved for designation of the most serious crimes by the FBI.
+In a way this could even mean the null values are also valuable, because they very well may indicate less serious crimes.
+
+```{r}
+d2<- d1 %>% 
+  mutate(
+    Family.Violence = toupper(Family.Violence)
+    ,Location.Type = if_else(
+      Location.Type == '',
+      'Unknown', Location.Type
+      )
+    ) %>% 
+  filter(!is.na(Zip.Code) & !(Zip.Code == 0) & !is.na(X.coordinate) & !is.na(Y.coordinate) & PRA != '',Clearance.Status != 9)
+skim(d2)
+print(count(d2)/count(d1))
+```
+
+The categorical variables look good for now.
+Up until this point little exploration has been done and there may be other re-codes or filters that would be worth including for other variables.
+Likely once I am further into the analysis I will return to this step to make more adjustments, but the changes I've made should provide a good starting point.
+
+### String Variables
+
+There are a few string variables that may be worth using for the analysis.
+Some of these I covered in the categorical variables, but for those that have excessive number of unique values a categorical value may not be practical.
+ - Highest Offense Description has hundreds of unique values, but since this is essentially a description of the crimes committed there are likely commonalities that could lead to groupings.
+I can imagine more variables stemming from this as well, such as if the description contains certain keywords I could create indicator variables based on that presence.
+ - Address is essentially another location based variable, and considering the number of them that have been cleaned for that purpose already it seems unlikely to need more.
+Still, the specificity offered by the address may be insightful, especially if i start drilling down into particular areas that suffer from the most crime.
+Potentially Doing something like a regular expression to extract a street name would be useful.
+ - Location Type does not have an unreasonable number of unique values, so this variable may better fit as a categorical only.
+Still the max length of some of the values lends itself to be that there may be more worth extracting from the field.
+
+Parsing out these string variables will likely be more of a cognitive load than the others would, so I will hold off on doing so for now.
+Once I am further along with the EDA I will have a better idea of what I want to be looking for within these variables anyway.
+
+### Tidied Up
+
+All done, data is clean now.
+
+Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.
+
+```{r}
+processeddata <- d2
+```
+
+# Save data
+
+Finally, we save the clean data as RDS file.
+
+```{r}
+save_data_location <- here("data","processed-data","processed-crime.rds")
+saveRDS(processeddata, file = save_data_location)
+```
+
+```{r Sandbox, echo=FALSE}
+#rawdata %>% filter(Occurred.Date.Time == '')
+
+rawdata %>% mutate(
+  test = (str_sub(as.character(Occurred.Time), -2, -1)
+          %>% as.numeric()),
+  Occurred.Date = mdy(Occurred.Date)
+  ,Report.Date = mdy(Report.Date)
+  ,Occurred.Date.Time = parse_date_time(Occurred.Date.Time,'%m/%d/%Y %I:%M:%S %p')
+  ,Report.Date.Time = parse_date_time(Report.Date.Time,'%m/%d/%Y %I:%M:%S %p')
+  ,Clearance.Date = mdy(Clearance.Date)
   ) %>% mutate(
@@
 ```
-
-Now we see that there is one person with a height of 6.
-That could be a typo, or someone mistakenly entered their height in feet.
-Since we unfortunately don't know, we might need to remove this person, which we'll do here.
-
-```{r}
-d2 <- d1 %>% dplyr::mutate( Height = replace(Height, Height=="6",round(6*30.48,0)) )
-skimr::skim(d2)
-```
-
-Height values seem ok now.
-
-Now let's look at the `Weight` variable.
-There is a person with weight of 7000, which is impossible, and one person with missing weight.
-To be able to analyze the data, we'll remove those individuals as well.
-
-```{r}
-d3 <- d2 %>%  dplyr::filter(Weight != 7000) %>% tidyr::drop_na()
-skimr::skim(d3)
-```
-
-Now checking the `Gender` variable.
-Gender should be a categorical/factor variable but is loaded as character.
-We can fix that with simple base R code to mix things up.
-
-```{r}
-d3$Gender <- as.factor(d3$Gender)  
-skimr::skim(d3)
-```
-
-Now we see that there is another NA, but it's not `NA` from R, instead it was loaded as character and is now considered as a category.
-Well proceed here by removing that individual with that NA entry.
-Since this keeps an empty category for Gender, I'm also using droplevels() to get rid of it.
-
-```{r}
-d4 <- d3 %>% dplyr::filter( !(Gender %in% c("NA","N")) ) %>% droplevels()
-skimr::skim(d4)
-```
-
-All done, data is clean now.
-
-Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.
-
-```{r}
-processeddata <- d4
-```
-
-# Save data
-
-Finally, we save the clean data as RDS file.
-I suggest you save your processed and cleaned data as RDS or RDA/Rdata files.
-This preserves coding like factors, characters, numeric, etc.
-If you save as CSV, that information would get lost.
-However, CSV is better for sharing with others since it's plain text.
-If you do CSV, you might want to write down somewhere what each variable is.
-
-See here for some suggestions on how to store your processed data: http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata
-
-```{r}
-save_data_location <- here::here("data","processed-data","processeddata.rds")
-saveRDS(processeddata, file = save_data_location)
-```
-
-# Notes
-
-Removing anyone who had "faulty" or missing data is one approach.
-It's often not the best.
-based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).
-
-```{r Sandbox}
-#rawdata %>% filter(Occurred.Date.Time == '')
-
-d1 <- rawdata %>% mutate(
-  test = (str_sub(as.character(Occurred.Time), -2, -1)
-          %>% as.numeric()),
-  Occurred.Date = mdy(Occurred.Date)
-  ,Report.Date = mdy(Report.Date)
-  ,Occurred.Date.Time = parse_date_time(Occurred.Date.Time,'%m/%d/%Y %I:%M:%S %p')
-  ,Report.Date.Time = parse_date_time(Report.Date.Time,'%m/%d/%Y %I:%M:%S %p')
-  ) %>% mutate(
-    Occurred.Time = if_else(
-      is.na(Occurred.Time) | test >= 60, 
-      if_else(
-        Occurred.Date == Report.Date,
-        if_else(
-          is.na(Report.Time), 
-          2359, Report.Time
-          ), 2359
-        ), if_else(
-          Occurred.Time == 2400, 
-          2359, Occurred.Time
-          ) 
-      ) %>% as.character()
-    ) %>% mutate(
-      Occurred.Time = paste0('000',Occurred.Time,'00') %>% 
-        str_sub(.,-6,-1) %>% 
-        parse_time('%H%M%S')
-      ) %>% select(-test)
-```

### 2024-06-22 products/manuscript/Manuscript.qmd
Minor changes made.
Lots of exploratory work done, deep dive into time series data, and many of the categorical variables in eda code. Moved a lot of that work out of the manuscript. Added descriptions, ideas, and interpretations. Started working on an ARIMA model for crime frequency.

--- 2024-06-14 products/manuscript/Manuscript.qmd
@@
 # load a few R packages
-library(here)
-library(knitr)
+pacman::p_load(here, knitr, tidyverse, skimr)
+theme_set(theme_minimal())
 ```
@@
 d1 <- readRDS(here('data','processed-data','processedcrime.rds'))
+zips <- read_csv(here('data','raw-data','austin-zip-codes.csv'))
 ```
@@
 
+```{r sandbox}
+ggplot() +
+    geom_sf(aes(fill = zip.code), data = zips2, inherit.aes = F, size = 0, alpha = 0.6, geom = zips2$the_geom) +
+    coord_sf(ndiscr = F) +
+    theme(legend.position = "none")
+```
+
 {{< pagebreak >}}

### 2024-06-22 code/analysis-code/statistical-analysis.R

@@ -1,3 +1,0 @@
-version https://git-lfs.github.com/spec/v1
-oid sha256:2e185b3e8e00e0617e772d144729b4161ff6784d1f63d08966bb2d951bf473ac
-size 1512

### 2024-06-22 code/eda-code/edacode.R

@@ -1,3 +1,0 @@
-version https://git-lfs.github.com/spec/v1
-oid sha256:0b112866218719c0d0f5936d491afb5c64f891b611f3d518c5d353b422a0e86f
-size 1432

### 2024-06-27 code/eda-code/eda.qmd
I think this was just minor updates to eda file. I'm only just noticing i'm behind a commit/push.

@@
          ) %>% 
-  filter(years == '2023' | years == '2022') %>% 
+  filter(years == '2003' | years == '2022') %>% 
   group_by(Family.Violence, Zip.Code, years) %>% 

### 2024-08-03 code/analysis-code/analysisfile.qmd
Project Assignment 4 and 5 submission. Mostly ready for final submission, need to add narrative to eda and stat qmd files

---
title: "Analysis"
author: "William Hyltin"
date: today
output: html_document
editor: 
  markdown: 
    wrap: sentence
---

# Setup

```{r setup}
#load needed packages. make sure they are installed.
pacman::p_load(here, knitr, tidyverse, skimr, fpp3, urca, forecast, gt)
theme_set(theme_minimal())
```

```{r}
d1 <- readRDS(here('data','processed-data','processed-crime.rds'))
```

```{r sandbox}

d1 %>% mutate(Category.Description = if_else(Category.Description == '', 'NA/Unknown', Category.Description)) %>% group_by(Category.Description, Highest.Offense.Description) %>% summarize(cnt = n()) %>% pivot_wider(names_from = Category.Description, values_from = cnt)

temp <- d1 %>%
         mutate(Occurred.Date = trunc.Date(Occurred.Date, 'months')) %>% 
         group_by(Occurred.Date) %>% 
         summarize(
             cnt = n()
         ) %>% as.data.frame() %>% 
  filter(year(Occurred.Date) != 2024)
rownames(temp) <- temp$Occurred.Date
temp <- temp %>% select(cnt)

temp2 <- temp %>% as.ts()
Acf(temp2)
Box.test(temp2, lag = 12, type = 'Ljung-Box')
auto.arima(temp2)
auto.arima(temp2) %>% forecast(h=24) %>% autoplot()
#temp2 %>% log() %>% diff(12) %>% diff() %>% diff() %>% diff() %>% diff() %>% diff() %>% Acf()
```

```{r}
temp3 <- d1 %>%
         filter(Crime.Category == 'MURDER') %>% 
         mutate(Occurred.Date = trunc.Date(Occurred.Date, 'months')) %>% 
         group_by(Occurred.Date) %>% 
         summarize(
             cnt = n()
         ) %>% as.data.frame() %>% 
  filter(year(Occurred.Date) < 2024)
rownames(temp3) <- temp3$Occurred.Date
temp3 <- temp3 %>% select(cnt)

temp5 <- temp3 %>% as.ts()
Acf(temp5)
auto.arima(log(temp5))
Box.test(diff(log(temp5)), lag = 4, type = 'Ljung-Box')
temp5 %>% log() %>% diff() %>% ur.kpss() %>% summary()

temp5 %>% log() %>% diff() %>% Acf()
temp5 %>% log() %>% diff() %>% plot()
```

```{r}
fit <- Arima(temp5, order = c(0,1,1))
autoplot(forecast(fit, h=24))
fit2 <- Arima(log(temp5), order = c(0,1,1))
autoplot(forecast(fit2, h=24))
```

```{r}
fets <- function(x, h) {
  forecast(ets(x), h=h)
}
farima <- function(x, h) {
  forecast(auto.arima(x), h=h)
}

e1 <- tsCV(log(temp5), fets, h=1)
e2 <- tsCV(log(temp5), farima, h=1)

# MSE comparisons
mean(e1^2, na.rm=TRUE)
mean(e2^2, na.rm=TRUE)
```

```{r}
crime_agg <- d1 %>%
  filter(Occurred.Date <= '2024-03-31') %>% 
  mutate(Occurred.Date = yearmonth(Occurred.Date)) %>% 
  group_by(Occurred.Date, Crime.Category) %>% 
  summarize(
     cnt = n()
  ) %>% as.data.frame() %>% 
  as_tsibble(key = Crime.Category, index = Occurred.Date) %>% 
  aggregate_key(Crime.Category, Count = sum(cnt)) %>% 
  fill_gaps(Count = 0)

group_fit <- crime_agg %>% 
  filter(year(Occurred.Date) < 2024) %>%
  model(base = ARIMA(Count)) %>% 
  reconcile(
    td = top_down(base)
  )

group_fit

group_fcast <- group_fit %>% 
  forecast(h = '2 years')

grpPlot <- group_fcast %>% 
  filter(Crime.Category %in% c('MURDER', 'AGGRAVATED ASSAULT', 'UNLAWFUL RESTRAINT' )| is_aggregated(Crime.Category), .model == 'td') %>% 
  autoplot(
    crime_agg %>% filter(year(Occurred.Date) >= 2019)#, Crime.Category == 'MURDER')
  ) + facet_wrap(vars(Crime.Category), scales = 'free_y',
                 labeller = labeller(
                   Crime.Category = c('<aggregated>' = 'All Crime', 
                                      'AGGRAVATED ASSAULT' = 'Aggravated Assault', 
                                      'MURDER' = 'Murder', 
                                      'UNLAWFUL RESTRAINT' = 'Unlawful Restraint')
                   )
                 ) +
  labs(x = 'Month', y = 'Count', title = 'Grouped Forecasts of Crime Occurence by Month') +
  theme(panel.background = element_rect(fill = 'white',
                                        color = 'white'),
        plot.background = element_rect(fill = 'white',
                                        color = 'white')
        )

murdplot <- group_fcast %>% 
  filter(Crime.Category %in% c('MURDER'), .model == 'td') %>% 
  autoplot(
    crime_agg %>% filter(year(Occurred.Date) >= 2019, Crime.Category == 'MURDER')
  ) +
  labs(x = 'Month', y = 'Count', title = 'Murder Forecasts by Month') +
  theme(panel.background = element_rect(fill = 'white',
                                        color = 'white'),
        plot.background = element_rect(fill = 'white',
                                        color = 'white')
        )

grpPlot  
murdplot

ggsave(here('results', 'figures', 'static-plots', 'crime-forecast.png'), grpPlot)
ggsave(here('results', 'figures', 'static-plots', 'murder-forecast.png'), murdplot)
```
```{r}
crime_agg %>% 
    filter(year(Occurred.Date) < 2024,Crime.Category == 'MURDER') %>%
    model(base = ARIMA(Count))

crime_agg %>% 
    filter(year(Occurred.Date) < 2024,Crime.Category == 'MURDER') %>%
    model(base = ARIMA(Count)) %>% forecast(h = '2 years') %>% 
  autoplot(crime_agg %>% filter(year(Occurred.Date) >= 2019, Crime.Category == 'MURDER'))
```

```{r}
AvF <- d1 %>% 
  filter(year(Occurred.Date) == 2024, Crime.Category == 'MURDER') %>% 
  mutate(
    Occurred.Date = yearmonth(Occurred.Date)
    ) %>% 
  group_by(Occurred.Date) %>% 
  summarize(
    n = n()
    ) %>% left_join(filter(group_fcast, Crime.Category == 'MURDER' & .model == 'td'), join_by(Occurred.Date == Occurred.Date)) %>% 
  select(Occurred.Date, n, .mean) %>% 
  gt() %>% 
  tab_header(
    title = md('**Forecast vs Actual<br>Murder Count**')
    ) %>% 
  cols_label(
    Occurred.Date = md('**Month**'),
    n = md('**Actual**'),
    .mean = md('**Forecasted**')
  ) %>% 
  fmt_number(columns = .mean, decimals = 1)

AvF

gtsave(AvF, 'actual-vs-forecast.html', here('results', 'figures', 'static-plots'))
```

```{r}
aggresid <- group_fit %>% filter(is_aggregated(Crime.Category)) %>% select(td) %>% gg_tsresiduals()

aggresid
ggsave(here('results', 'figures', 'static-plots', 'crime-resid.png'), aggresid)
```

```{r}
murdresid <- group_fit %>% filter(Crime.Category == 'MURDER') %>% select(td) %>% gg_tsresiduals()

murdresid
ggsave(here('results', 'figures', 'static-plots', 'murd-resid.png'), murdresid)
```

```{r}
agglb <- augment(filter(group_fit, is_aggregated(Crime.Category))) %>% 
  features(.resid, ljung_box, l = 24) %>% gt()

agglb
gtsave(agglb, 'aggregated-ljung.html', here('results', 'figures', 'static-plots'))
```

```{r}
murdlb <- augment(filter(group_fit, Crime.Category == 'MURDER')) %>% 
  features(.resid, ljung_box, l = 24) %>% gt()

murdlb
gtsave(murdlb, 'murder-ljung.html', here('results', 'figures', 'static-plots'))
```

### 2024-08-03 code/eda-code/eda.qmd
Project Assignment 4 and 5 submission. Mostly ready for final submission, need to add narrative to eda and stat qmd files

@@
 title: "An example exploratory analysis script"
-date: "2024-02-07"
+date: "2024-07-31"
 output: html_document
@@
 #load needed packages. make sure they are installed.
-pacman::p_load(here, knitr, tidyverse, skimr, fpp2)
+pacman::p_load(here, knitr, tidyverse, skimr, fpp2, tigris, plotly, viridis, transformr, htmlwidgets, gt, gtExtras, widgetframe)
 theme_set(theme_minimal())
@@
 d1 <- readRDS(here('data','processed-data','processed-crime.rds'))
-zips <- read_csv(here('data','raw-data','austin-zip-codes.csv'))
+tig_zips <- zctas(cb=TRUE, starts_with = c(unique(d1$Zip.Code)), year = 2020)
 ```
@@
 ```{r}
-summary(zips)
-head(zips)
-str(zips)
-skim(zips)
+summary(tig_zips)
+head(tig_zips)
+str(tig_zips)
+skim(tig_zips)
 ```
@@
 ```{r}
-d1 %>% group_by(Occurred.Date) %>% 
+dayline <- d1 %>% group_by(Occurred.Date) %>% 
   summarize(
@@
   ggplot(mapping = aes(x=Occurred.Date, y=cnt)) +
-  geom_line()
+  geom_line() +
+  labs(title = 'Austin Crime Over Time', x = 'Time of Occurence (by day)', y = 'Count')
+
+dayline
+
+ggsave(here('results', 'figures', 'static-plots', 'crime-by-day.png'), dayline)
 ```
@@
 ```{r}
-d1 %>% filter(Occurred.Date <= '2023-12-31') %>%
+monthline <- d1 %>% filter(Occurred.Date <= '2024-03-31') %>%
   mutate(Occurred.Date = trunc.Date(Occurred.Date, 'months')) %>% 
@@
   ggplot(mapping = aes(x=Occurred.Date, y=cnt)) +
-  geom_line()
+  geom_line() +
+  labs(title = 'Austin Crime Over Time', x = 'Time of Occurence (by month)', y = 'Count')
+
+monthline
+
+ggsave(here('results', 'figures', 'static-plots', 'crime-by-month.png'), monthline)
 ```
@@
 This may be another one to layer in with time, like a chart with a slider for the year to see shifts in crime frequency by month over time.
+
+```{r}
+murdline <- d1 %>% filter(Occurred.Date <= '2024-03-31', Crime.Category == 'MURDER') %>%
+  mutate(Occurred.Date = trunc.Date(Occurred.Date, 'months')) %>% 
+  group_by(Occurred.Date) %>% 
+  summarize(
+  cnt = n()
+) %>% as.data.frame() %>% 
+  ggplot(mapping = aes(x=Occurred.Date, y=cnt)) +
+  geom_line() +
+  labs(title = 'Austin Murder Over Time', x = 'Time of Occurence (by month)', y = 'Count')
+
+murdline
+
+ggsave(here('results', 'figures', 'static-plots', 'murder-by-time.png'), murdline)
+```
+*ADD COMMMENTARY LATER: NUMBER OF FIRST QUARTER HOMICIDES MATCHES ARTICLE FOR 2020 AND 2023, BUT NOT 2024.
+ARTICLE LIKELY NOT COUNTING INTOX MANSLAUGHTER OR JUSTIFIED HOMICIDE.
+KEPT THOSE BECAUSE FBI CATEGORIZATION COUNTS MANSLAUGHTER.*
 
@@
 
-# Notes
+```{r}
+yearlyCrime <- d1 %>% mutate(
+  Zip.Char = as.character(Zip.Code),
+  years = year(Occurred.Date),
+  years = as.character(years)
+  ) %>% filter(years != 2024) %>% 
+  group_by(Zip.Char, years) %>% 
+  summarize(crim.count  = n()) %>% 
+  ungroup() %>% 
+  as.data.frame()
+  #filter(count >= 500) %>%
+  #filter(Zip.Char == '78741') %>%
 
-Including this sandbox chunk where I can experiment with code as needed.
-ARIMA is not going well, mostly due to seasonal nature of data. Filtering the time range might help, or isolating other variables to do things liek predict certain types of crime instead.
+allYrsZips <- expand.grid(years = unique(yearlyCrime$years), Zip.Char = unique(yearlyCrime$Zip.Char))
 
-```{r sandbox}
+yearlyCrimeAll <- left_join(allYrsZips, yearlyCrime, by = join_by(Zip.Char, years))  %>% 
+  inner_join(tig_zips, by = join_by(Zip.Char == ZCTA5CE20)) %>% 
+  mutate(crim.count = ifelse(is.na(crim.count), 0, crim.count))
 
-d1 %>% mutate(Category.Description = if_else(Category.Description == '', 'NA/Unknown', Category.Description)) %>% group_by(Category.Description, Highest.Offense.Description) %>% summarize(cnt = n()) %>% pivot_wider(names_from = Category.Description, values_from = cnt)
+```
 
-temp <- d1 %>%
-         mutate(Occurred.Date = trunc.Date(Occurred.Date, 'months')) %>% 
-         group_by(Occurred.Date) %>% 
+```{r}
+cpeth <- yearlyCrimeAll %>% filter(crim.count >= 50) %>% 
+  ggplot() +
+  geom_sf(aes(geometry = geometry, fill = crim.count, frame = years, text = paste(Zip.Char, '<br>', crim.count))) +
+  scale_fill_gradient2(low = "#E0144C", mid = '#FFFFFF', high = "#000067", midpoint = -5000, 
+                       trans = 'reverse') +
+  theme_classic() +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.text.y=element_blank(),
+        line = element_blank(),
+        plot.title = element_text(hjust = 0, size = 15)) +
+  labs(fill = 'Crime Count',
+       title = 'Crime by Location Over Time')
+
+cplotly <- cpeth %>% ggplotly() %>% 
+  layout(hoverlabel = text) %>% 
+  animation_opts(1500, 1) %>% 
+  animation_slider(currentvalue = list(visible = FALSE)) %>% 
+  animation_button(x = 0, xanchor = "left", y = 0, yanchor = "bottom")
+cplotly
+
+saveWidget(frameableWidget(cplotly), here('results', 'figures', 'html-widgets', 'crime-location.html'))
+```
+
+```{r}
+yearlyMurder <- d1 %>% mutate(
+  Zip.Char = as.character(Zip.Code),
+  years = year(Occurred.Date),
+  years = as.character(years)
+  ) %>% filter(years != 2024, Crime.Category == 'MURDER') %>% 
+  group_by(Zip.Char, years) %>% 
+  summarize(murder.count  = n()) %>% 
+  ungroup() %>% 
+  as.data.frame()
+
+yearlyMurderAll <- left_join(allYrsZips, yearlyMurder, by = join_by(Zip.Char, years))  %>% 
+  inner_join(tig_zips, by = join_by(Zip.Char == ZCTA5CE20)) %>% 
+  mutate(murder.count = ifelse(is.na(murder.count), 0, murder.count))
+```
+
+```{r}
+cpeth2 <- yearlyMurderAll %>% #filter(murder.count >= 1) %>% 
+  ggplot() +
+  geom_sf(aes(geometry = geometry, fill = murder.count, frame = years)) +
+  #geom_sf_label(aes(label = paste(Zip.Char, '<br>', murder.count))) +
+  scale_fill_gradient2(low = "#E0144C", mid = '#FFFFFF', high = "#000067", 
+                       trans = 'reverse') +
+  theme_classic() +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.text.y=element_blank(),
+        line = element_blank(),
+        plot.title = element_text(hjust = 0, size = 15)) +
+  labs(fill = 'Murder Count',
+       title = 'Murder by Location Over Time')
+
+cplotly2 <- cpeth2 %>% ggplotly() %>% 
+  layout(hoverlabel = text) %>% 
+  animation_opts(1500, 1) %>% 
+  animation_slider(currentvalue = list(visible = FALSE)) %>% 
+  animation_button(x = 0, xanchor = "left", y = 0, yanchor = "bottom")
+cplotly2
+```
+
+```{r}
+catYearly <- d1 %>% mutate(
+  years = year(Occurred.Date)
+  #years = as.character(years)
+  ) %>% filter(years != 2024) %>% 
+  group_by(Crime.Category, years) %>% 
+  summarize(crim.count  = n()) %>% 
+  ungroup() %>% 
+  as.data.frame()
+
+allYrsCats <- expand.grid(years = unique(catYearly$years), Crime.Category = unique(catYearly$Crime.Category))
+
+yearlyAllCats <- left_join(allYrsCats, catYearly, by = join_by(Crime.Category, years))  %>% 
+  mutate(crim.count = ifelse(is.na(crim.count), 0, crim.count))
+
+catbar <- yearlyAllCats %>%  
+  ggplot() +
+  geom_bar(aes(y= fct_reorder(Crime.Category, crim.count), x = crim.count, frame = years), stat = 'identity') +
+  labs(x = 'Count of Offense', y = 'Offense Description', title = 'Frequncy of Crim by Offense Category')
+  #scale_x_continuous(transform = 'log')
+catbar
+
+fig <- plot_ly(yearlyAllCats,
+  x = ~crim.count,
+  y = ~fct_reorder(Crime.Category, crim.count),
+  type = 'bar',
+  showlegend = F,
+  frame = ~years,
+  marker = list(color = '#000067')
+)
+
+catbarly <- fig %>% 
+  layout(
+    yaxis = list(
+      title = '', tickangle = -30, tickfont = list(size = 8)
+      ), 
+    xaxis = list(
+      title = 'Count of Occurence' 
+    ),
+    title = list(
+      text = 'Yearly Occurence of Crime by Category',
+      y = 0.99, x = 0.1, xanchor = 'left', yanchor = 'top',
+      font = list(
+        size = 18
+      )
+      )
+    )
+
+catbarly
+
+saveWidget(frameableWidget(catbarly), here('results', 'figures', 'html-widgets', 'crime-cat.html'))
+```
+
+First table is intended to act as a translation key between description and category, but also serves well to illustrate the motivation behind the categorizations.
+```{r}
+RecatTable <- d1 %>% 
+  filter(year(Occurred.Date) < 2024) %>% 
+  mutate(
+  Occur.Years = trunc.Date(Occurred.Date, 'years')
+  ) %>% 
+  group_by(Crime.Category, Highest.Offense.Description, Occur.Years) %>% 
   summarize(
-             cnt = n()
-         ) %>% as.data.frame() %>% 
-  filter(year(Occurred.Date) != 2024)
-rownames(temp) <- temp$Occurred.Date
-temp <- temp %>% select(cnt)
+    Count = n(),
+    .groups = 'drop'
+  ) %>% 
+  arrange(by = Occur.Years) %>%
+  group_by(Crime.Category, Highest.Offense.Description) %>% 
+  summarize(
+    mean = mean(Count),
+    sd = sd(Count),
+    sum = sum(Count),
+    .groups = 'drop'
+  ) %>% 
+  gt() %>% 
+  tab_header(
+    title = md('**Crime Recategorization Key**')
+    ) %>% 
+  cols_label(
+    Crime.Category = md('**New Category**'),
+    Highest.Offense.Description = md('**Original Description**'),
+    mean = md('**Mean Occurrence Across Years**'),
+    sd = md('**Standard Deviation of Yearly Occurrence**'),
+    sum = md('**Total Occurrences**')
+  ) %>% 
+  fmt_number(columns = mean:sd, decimals = 1) %>% 
+  opt_interactive()
 
-temp2 <- temp %>% as.ts()
-Acf(temp2)
-Box.test(temp2, lag = 12, type = 'Ljung-Box')
-auto.arima(temp2)
-temp2 %>% log() %>% diff(12) %>% diff() %>% diff() %>% diff() %>% diff() %>% diff() %>% Acf()
+gtsave(RecatTable, 'recat-table.html', here('results', 'figures', 'html-widgets'))
+#saveWidget(frameableWidget(RecatTable), here('results', 'figures', 'html-widgets', 'recat-table.html'))
+RecatTable
 ```
+
+Plot highlights the advantages in the categorization, with inclusion of sparklines for clarity on movement over time.
+```{r}
+newCatTable <- d1 %>%
+  filter(year(Occurred.Date) < 2024) %>% 
+  mutate(
+  Occur.Years = trunc.Date(Occurred.Date, 'years')
+  ) %>% 
+  group_by(Crime.Category, Occur.Years) %>% 
+  summarize(
+    Count = n(),
+    .groups = 'drop'
+  ) %>% 
+  arrange(by = Occur.Years) %>%
+  group_by(Crime.Category) %>% 
+  summarize(
+    mean = mean(Count),
+    sd = sd(Count),
+    list = list(Count),
+    .groups = 'drop'
+  ) %>% 
+  gt() %>% 
+  tab_header(
+    title = md('**Crime Categories Over Time**')
+    ) %>% 
+  cols_label(
+    Crime.Category = md('**Crime Category**'),
+    mean = md('**Mean Occurrence Across Years**'),
+    sd = md('**Standard Deviation of Yearly Occurrence**'),
+    list = md('**Change Over Time**')
+  ) %>% 
+  fmt_number(columns = mean:sd, decimals = 1) %>% 
+  gtExtras::gt_plt_sparkline(column = list, type = 'ref_iqr', same_limit = FALSE) %>% 
+  tab_footnote('Note: Sparklines have independent axes, comparisons should not be considered across different categories')
+gtsave(newCatTable, 'new-cat-table.html', here('results', 'figures', 'html-widgets'))
+newCatTable
+
+```

### 2024-08-03 code/processing-code/processingfile.qmd
Project Assignment 4 and 5 submission. Mostly ready for final submission, need to add narrative to eda and stat qmd files

@@
 
+### And We're Back
+
+The string variable `Highest.Offense.Description` was just too rich to pass up.
+It would make for an excellent categorical variable however there are over 400 categories going off the variable as-is.
+`UCR.Category` is a very nice alternative with only a few levels to the category, but the field is mostly empty due to the nature of the variable.
+It represent an FBI Categorization known as Uniform Crime Reporting (UCR) for the most serious of crimes.
+So I decided to take the matter into my own hands.
+Number one, I would create a new field that categorizes the crime myself, based on categorizations in the UCR program.
+If the observation had values in the UCR field, I would take that one, if it did not, I impute a category based on the description in the Highest Offense Description field.
+This was mostly done through brute force, with string detection and research in two places, the [FBI UCR archive](https://ucr.fbi.gov/crime-in-the-u.s/2019/crime-in-the-u.s.-2019/topic-pages/offense-definitions) (2019 was not the only year that had information on the categorizations but it was the most in a quick reference format), and the [Texas Capitol Statutes website](https://statutes.capitol.texas.gov/).
+I feel the need to address this up front, I am absolutely not a criminologist, these categorizations are solely for the sake of making my analysis easier.
+I think of this as grouping the crime descriptions from the perspective of a layman, me, because ultimately the average person observing crime statistics will roughly as little knowledge about them as I do.
+An expert could and would do a much better job than I did, but that wasn't readily available so I worked with what I had.
+Famous last words of an analyst.
+```{r}
+unique(d2$Category.Description) %>% as.data.frame()
+unique(d2[c('Category.Description','Highest.Offense.Description')]) %>% as.data.frame()
+```
+
+```{r}
+UCRCats <- d2 %>% filter(Category.Description != '') %>% select(Category.Description) %>% unique()
+row.names(UCRCats) <- NULL
+UCRCats <- UCRCats[c(7,1,4,5,6,3,2),]
+row.names(UCRCats) <- NULL
+
+defCats <- rbind(CHILD = str_flatten(c('CHILD', 'MINOR', 'TRUANCY', 'NONSUPPORT'), collapse = '|') # Sex Offense involving a Child, Child Abuse/ Endangerment, Injury/Harm of Vulnerable Persons*
+                 ,HARASSMENT = str_flatten(c('HARASSMENT', 'PROWLER', 'STALKING'), collapse = '|') # General Harassment
+                 ,'SEX OFFENSE'= str_flatten(c('SEX', 'INVASIVE', 'INTIMATE', 'PROSTITUTION', 'OBSC', 'ADULT', 'BESTIALITY', 'INDECEN', 'ENTIC', 'PORN', 'SOLICITATION OF', 'GROOMING'), collapse = '|') # Sexual Offense
+                 ,'DISORDERLY CONDUCT' = str_flatten(c('DOC', 'LEWD', 'URINATING', 'RECKLESS CONDUCT', 'ILLUMIN', 'CORPSE', 'ANIMAL', 'RIOT', 'PROCESSION'), collapse = '|') #Disorderly Conduct
+                 ,VOCO = str_flatten(c('VOCO', 'ORDINANCE', 'CURFEW', 'CAMP', 'AMPLIFIED', 'GLASS', 'CITY ORD'), collapse = '|') # Violation of City Ordinance
+                 ,'SIMPLE ASSAULT' = str_flatten(c('ASSAULT','ASLT$', 'ASSLT', 'PROD', 'ATTACK BY DOG', 'VIOLENCE', 'HAZING', 'SUICIDE', 'CONFRONT'), collapse = '|') # Simple Assault/ ASSAULTIVE
+                 ,TRESPASSING = str_flatten(c('TRESPASS', 'LOITER', 'SLEEP', 'BEG'), collapse = '|') # TRESPASSING
+                 ,'LARGE SCALE THREAT' = str_flatten(c('THREAT'), collapse = '|') # Large Scale Threat
+                 ,'GENERAL DISTURBANCE' = str_flatten(c('DISTURBANCE'), collapse = '|') # General Disturbance
+                 ,'DRUG RELATED' = str_flatten(c('DRUG', 'SUB', 'MARIJUANA', 'INHALANT', 'TOBACCO', 'PRESCRIPTION'), collapse = '|') #Drug Related
+                 ,'FRAUD' = str_flatten(c('FRAUD', 'FALSE', 'IMPERSON', 'FORGE', 'COUNTERFEITING', 'TAMPER', 'LABEL', 'PIGEON', 'MEDICINE', 'NAME'), collapse = '|') #Fraud, Unlawful Deception
+                 ,'ALCOHOL RELATED' = str_flatten(c('ALCOHOL', 'LIQ', 'DWI', 'DUI', 'WHILE INTOX', 'BWI'), collapse = '|') # Alcohol Related
+                 ,'HARM OF VULNERABLE PERSONS' = str_flatten(c('INJ', 'ELDER'), collapse = '|') #Injury/Harm of Vulnerable Persons*
+                 ,'NON-COMPLIANCE' = str_flatten(c('WARR', 'BAIL', 'IDENTIFY', 'PROBATION', 'INTER', '911', 'ARREST', 'EVAD', 'PERJURY', 'PAROLE','RETALIATION', 'ESCAPE', 'HINDERING', 'BOND', 'CONTEMPT'), collapse = '|') # Warrant, Non Compliance
+                 ,'CRIMINAL MISCHIEF' = str_flatten(c('MISCHIEF', 'ARSON', 'DAMAGE', 'GRAFFITI', 'DESECRATION'), collapse = '|') # CRIMINAL MISCHIEF
+                 ,'UNLAWFUL RESTRAINT' = str_flatten(c('KIDNAP', 'RESTRAINT'), collapse = '|') # Unlawful Restraint
+                 ,'WEAPON RELATED' = str_flatten(c('WEAPON', 'FIREARM', 'EXPLOSIVE', 'UCW', 'KNIFE', 'CARRY', 'HANDGUN'), collapse = '|') # WEAPON RELATED
+                 ,'PROTECTIVE ORDER' = str_flatten(c('ORDER'), collapse = '|') #PROTECTIVE/ RESTRAINING ORDER
+                 ,'FINANCIAL CRIME' = str_flatten(c('CRED', 'DEBIT', 'MONEY', 'BANK'), collapse = '|') # FINANCIAL CRIME
+                 ,LITTERING = str_flatten(c('LITTER', 'DUMP', 'WASTE', 'REFRIGERATOR'), collapse = '|') #LITTERING
+                 ,'CRIMINAL CONSPIRACY' = str_flatten(c('CRIMINAL SOLICITATION', 'ABATE', 'ORGANIZED', 'CONSPIR'), collapse = '|') # CRIMINAL CONSPIRACY
+                 ,'ABUSE OF OFFICE' = str_flatten(c('OFFICIAL', 'PRISONER'), collapse = '|') # ABUSE OF OFFICE
+                 ,'GAMBLING RELATED' = str_flatten(c('GAMBL'), collapse = '|') # GAMBLING RELATED
+                 ,THEFT = str_flatten(c('RENTAL CAR'), collapse = '|')# THEFT
+                 ,MURDER = str_flatten(c('MANSLAUGHTER', 'HOMICIDE'), collapse = '|') # MURDER (MANSLAUGHTER INCLUDED BECAUSE THAT'S HOW FBI CATS)
+                 ,BRIBERY = str_flatten(c('BRIBE', 'COERCION OF PUBLIC SERVANT', 'GIFT'), collapse = '|') # BRIBERY
+                 ,OTHER = str_flatten(c('IMMIGRATION', 'ALIEN', 'OTHER', 'DISRUPTION OF CLASSES', 'AIRPORT', 'INSTRUMENT','SCHOOLS', 'MASSAGE', 'WATER', 'AWOL', 'DESERTION', 'RENDER'), collapse = '|')# OTHER
+                 )%>% as.data.frame()
+defCats$Cat <- rownames(defCats)
+rownames(defCats) <- NULL  
+colnames(defCats) <- c('strSearch', 'Cats')
+
+d3 <- d2 %>% 
+    mutate(Crime.Category = if_else(Category.Description == '', 
+                                    str_extract(Highest.Offense.Description, str_flatten(toupper(UCRCats), collapse = '|')), 
+                                    toupper(Category.Description)
+                                    )
+          ,Crime.Category = if_else(is.na(Crime.Category),
+                                    case_when(
+                                      str_detect(Highest.Offense.Description,defCats[1,1]) == TRUE ~ defCats[1,2],
+                                      str_detect(Highest.Offense.Description,defCats[2,1]) == TRUE ~ defCats[2,2],
+                                      str_detect(Highest.Offense.Description,defCats[3,1]) == TRUE ~ defCats[3,2],
+                                      str_detect(Highest.Offense.Description,defCats[4,1]) == TRUE ~ defCats[4,2],
+                                      str_detect(Highest.Offense.Description,defCats[5,1]) == TRUE ~ defCats[5,2],
+                                      str_detect(Highest.Offense.Description,defCats[6,1]) == TRUE ~ defCats[6,2],
+                                      str_detect(Highest.Offense.Description,defCats[7,1]) == TRUE ~ defCats[7,2],
+                                      str_detect(Highest.Offense.Description,defCats[8,1]) == TRUE ~ defCats[8,2],
+                                      str_detect(Highest.Offense.Description,defCats[9,1]) == TRUE ~ defCats[9,2],
+                                      str_detect(Highest.Offense.Description,defCats[10,1]) == TRUE ~ defCats[10,2],
+                                      str_detect(Highest.Offense.Description,defCats[11,1]) == TRUE ~ defCats[11,2],
+                                      str_detect(Highest.Offense.Description,defCats[12,1]) == TRUE ~ defCats[12,2],
+                                      str_detect(Highest.Offense.Description,defCats[13,1]) == TRUE ~ defCats[13,2],
+                                      str_detect(Highest.Offense.Description,defCats[14,1]) == TRUE ~ defCats[14,2],
+                                      str_detect(Highest.Offense.Description,defCats[15,1]) == TRUE ~ defCats[15,2],
+                                      str_detect(Highest.Offense.Description,defCats[16,1]) == TRUE ~ defCats[16,2],
+                                      str_detect(Highest.Offense.Description,defCats[17,1]) == TRUE ~ defCats[17,2],
+                                      str_detect(Highest.Offense.Description,defCats[18,1]) == TRUE ~ defCats[18,2],
+                                      str_detect(Highest.Offense.Description,defCats[19,1]) == TRUE ~ defCats[19,2],
+                                      str_detect(Highest.Offense.Description,defCats[20,1]) == TRUE ~ defCats[20,2],
+                                      str_detect(Highest.Offense.Description,defCats[21,1]) == TRUE ~ defCats[21,2],
+                                      str_detect(Highest.Offense.Description,defCats[22,1]) == TRUE ~ defCats[22,2],
+                                      str_detect(Highest.Offense.Description,defCats[23,1]) == TRUE ~ defCats[23,2],
+                                      str_detect(Highest.Offense.Description,defCats[24,1]) == TRUE ~ defCats[24,2],
+                                      str_detect(Highest.Offense.Description,defCats[25,1]) == TRUE ~ defCats[25,2],
+                                      str_detect(Highest.Offense.Description,defCats[26,1]) == TRUE ~ defCats[26,2],
+                                      str_detect(Highest.Offense.Description,defCats[27,1]) == TRUE ~ defCats[27,2],
+                                      .default = Crime.Category
+                                    ),
+                                    Crime.Category
+          )
+          ,Crime.Category = if_else(Crime.Category == 'CHILD',
+                                    case_when(
+                                      str_detect(Highest.Offense.Description,defCats[3,1]) == TRUE ~ 'SEX OFFENSE INVOLVING A CHILD',
+                                      .default = 'HARM OF VULNERABLE PERSONS'
+                                    ),
+                                    Crime.Category
+                                    )
+          ,Crime.Category = if_else(Highest.Offense.Description == 'JUSTIFIED HOMICIDE', 'JUSTIFIED HOMICIDE', Crime.Category)
+          )
+d3 %>% select(Highest.Offense.Description, Category.Description, Crime.Category) %>% head(100)
+d3 %>% dim()
+
+d3 %>% group_by(Crime.Category, Category.Description) %>% summarize(
+  cnt = n()
+)
+
+d3 %>% filter(!is.na(Crime.Category)) %>%
+  #,Category.Description == '') %>% 
+  #select(Crime.Category, Highest.Offense.Description) %>% 
+  group_by(Crime.Category) %>%  #, Highest.Offense.Description) %>% 
+  summarize(
+    cnt = n()
+  ) %>% 
+  arrange(by = cnt)
+```
+
 ### Tidied Up
@@
 ```{r}
-processeddata <- d2
+processeddata <- d3
 ```
@@
 ```
-
-```{r Sandbox, echo=FALSE}
-#rawdata %>% filter(Occurred.Date.Time == '')
-
-rawdata %>% mutate(
-  test = (str_sub(as.character(Occurred.Time), -2, -1)
-          %>% as.numeric()),
-  Occurred.Date = mdy(Occurred.Date)
-  ,Report.Date = mdy(Report.Date)
-  ,Occurred.Date.Time = parse_date_time(Occurred.Date.Time,'%m/%d/%Y %I:%M:%S %p')
-  ,Report.Date.Time = parse_date_time(Report.Date.Time,'%m/%d/%Y %I:%M:%S %p')
-  ,Clearance.Date = mdy(Clearance.Date)
-  ) %>% mutate(
-    Occurred.Time = if_else(
-      is.na(Occurred.Time) | test >= 60, 
-      if_else(
-        Occurred.Date == Report.Date,
-        if_else(
-          is.na(Report.Time), 
-          2359, Report.Time
-          ), 2359
-        ), if_else(
-          Occurred.Time == 2400, 
-          2359, Occurred.Time
-          ) 
-      ) %>% as.character()
-    ) %>% mutate(
-      Occurred.Time = paste0('000',Occurred.Time,'00') %>% 
-        str_sub(.,-6,-1) %>% 
-        parse_time('%H%M%S')
-      ) %>% select(-test)
-```

### 2024-08-03 products/manuscript/Manuscript.qmd
Project Assignment 4 and 5 submission. Mostly ready for final submission, need to add narrative to eda and stat qmd files

--- 2024-06-14 products/manuscript/Manuscript.qmd
@@
 format:
-  docx:
-    toc: false
-    number-sections: true
-    highlight-style: github
+  html:
+    embed-resources: true
+    code-fold: true
+editor: 
+  markdown: 
+    wrap: sentence
 bibliography: ../../assets/dataanalysis-references.bib
-csl: ../../assets/apa.csl
+csl: ../../assets/data-in-brief.csl
 ---
@@
 # load a few R packages
-library(here)
-library(knitr)
+pacman::p_load(here, knitr, tidyverse, skimr, htmlwidgets, htmltools, plotly, tigris, gt, gtExtras)
+theme_set(theme_minimal())
 ```
 
-# Summary/Abstract
-_Write a summary of your project._
+```{r Load Processed Data and Figures, message = FALSE, warning = FALSE}
+d1 <- readRDS(here('data','processed-data','processed-crime.rds'))
+tig_zips <- zctas(cb=TRUE, starts_with = c(unique(d1$Zip.Code)), year = 2020, progress_bar = FALSE)
 
-{{< pagebreak >}}
+catYearly <- d1 %>% mutate(
+  years = year(Occurred.Date)
+  #years = as.character(years)
+  ) %>% filter(years != 2024) %>% 
+  group_by(Crime.Category, years) %>% 
+  summarize(crim.count  = n()) %>% 
+  ungroup() %>% 
+  as.data.frame()
+
+allYrsCats <- expand.grid(years = unique(catYearly$years), Crime.Category = unique(catYearly$Crime.Category))
+
+yearlyAllCats <- left_join(allYrsCats, catYearly, by = join_by(Crime.Category, years))  %>% 
+  mutate(crim.count = ifelse(is.na(crim.count), 0, crim.count))
+
+fig <- plot_ly(yearlyAllCats,
+  x = ~crim.count,
+  y = ~fct_reorder(Crime.Category, crim.count),
+  type = 'bar',
+  showlegend = F,
+  frame = ~years,
+  marker = list(color = '#000067')
+)
+
+catbarly <- fig %>% 
+  layout(
+    yaxis = list(
+      title = '', tickangle = -30, tickfont = list(size = 8)
+      ), 
+    xaxis = list(
+      title = 'Count of Occurence' 
+    ),
+    title = list(
+      text = 'Yearly Occurence of Crime by Category',
+      y = 0.99, x = 0.1, xanchor = 'left', yanchor = 'top',
+      font = list(
+        size = 18
+      )
+      )
+    )
+
+yearlyCrime <- d1 %>% mutate(
+  Zip.Char = as.character(Zip.Code),
+  years = year(Occurred.Date),
+  years = as.character(years)
+  ) %>% filter(years != 2024) %>% 
+  group_by(Zip.Char, years) %>% 
+  summarize(crim.count  = n()) %>% 
+  ungroup() %>% 
+  as.data.frame()
+  #filter(count >= 500) %>%
+  #filter(Zip.Char == '78741') %>%
+
+allYrsZips <- expand.grid(years = unique(yearlyCrime$years), Zip.Char = unique(yearlyCrime$Zip.Char))
+
+yearlyCrimeAll <- left_join(allYrsZips, yearlyCrime, by = join_by(Zip.Char, years))  %>% 
+  inner_join(tig_zips, by = join_by(Zip.Char == ZCTA5CE20)) %>% 
+  mutate(crim.count = ifelse(is.na(crim.count), 0, crim.count))
+
+cpeth <- yearlyCrimeAll %>% filter(crim.count >= 50) %>% 
+  ggplot() +
+  geom_sf(aes(geometry = geometry, fill = crim.count, frame = years, text = paste(Zip.Char, '<br>', crim.count))) +
+  scale_fill_gradient2(low = "#E0144C", mid = '#FFFFFF', high = "#000067", midpoint = -5000, 
+                       trans = 'reverse') +
+  theme_classic() +
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.text.y=element_blank(),
+        line = element_blank(),
+        plot.title = element_text(hjust = 0, size = 15)) +
+  labs(fill = 'Crime Count',
+       title = 'Crime by Location Over Time')
+
+cplotly <- cpeth %>% ggplotly() %>% 
+  layout(hoverlabel = text) %>% 
+  animation_opts(1500, 1) %>% 
+  animation_slider(currentvalue = list(visible = FALSE)) %>% 
+  animation_button(x = 0, xanchor = "left", y = 0, yanchor = "bottom")
+
+RecatTable <- d1 %>% 
+  filter(year(Occurred.Date) < 2024) %>% 
+  mutate(
+  Occur.Years = trunc.Date(Occurred.Date, 'years')
+  ) %>% 
+  group_by(Crime.Category, Highest.Offense.Description, Occur.Years) %>% 
+  summarize(
+    Count = n(),
+    .groups = 'drop'
+  ) %>% 
+  arrange(by = Occur.Years) %>%
+  group_by(Crime.Category, Highest.Offense.Description) %>% 
+  summarize(
+    mean = mean(Count),
+    sd = sd(Count),
+    sum = sum(Count),
+    .groups = 'drop'
+  ) %>% 
+  gt() %>% 
+  tab_header(
+    title = md('**Crime Recategorization Key**')
+    ) %>% 
+  cols_label(
+    Crime.Category = md('**New Category**'),
+    Highest.Offense.Description = md('**Original Description**'),
+    mean = md('**Mean Occurrence Across Years**'),
+    sd = md('**Standard Deviation of Yearly Occurrence**'),
+    sum = md('**Total Occurrences**')
+  ) %>% 
+  fmt_number(columns = mean:sd, decimals = 1) %>% 
+  opt_interactive()
+```
 
@@
 ## General Background Information
-[Last year](https://www.fox7austin.com/news/austin-named-city-with-15th-biggest-homicide-rate-problem-study), Austin was ranked among the top US cities with a problem in homicide rate.
-Earlier [this year](https://www.kvue.com/article/news/local/austin-crime-rate-drops-2024-2020-comparison/269-b62042b7-a6f2-4224-87ff-ede655a37912), the claim was made that crime has dropped to a new low since 2021.
+
+[Last year](https://www.fox7austin.com/news/austin-named-city-with-15th-biggest-homicide-rate-problem-study), Austin was ranked among the top US cities with a problem in homicide rate [@fox2023].
+Earlier [this year](https://www.kvue.com/article/news/local/austin-crime-rate-drops-2024-2020-comparison/269-b62042b7-a6f2-4224-87ff-ede655a37912), the claim was made that crime has dropped to a new low since 2020 [@fox2024].
 The intention of this analysis will be to investigate both claims, and gain a better understanding of Austin's criminal activity.
@@
 I may identify shifts that could inform the cause, but the goal will not be to drill down into every potential socioeconomic factor and pick the primary drivers, instead simply to observe what crime in the city of Austin looks like and what we might be able to expect going forward.
-Aside from resources outlined in our course resources, I was hoping to also draw from this [book](https://otexts.com/fpp2/index.html).
+Aside from resources outlined in our course resources, I will also draw from this [book](https://otexts.com/fpp3/) [@hyndman2021].
 It's not one I have used before but it had been recommended to me previously and I would like to learn more about time series and forecasting in general.
@@
 ## Description of data and data source
-My primary data source will come from [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data), but I will likely need to pull from other sources as well, potentially finding other cities' crime stats that made the same list of problematic homicide rates, or even Austin's demographic information and statistics like population size.
 
-The raw data has several data and time fields, many of which are character strings meaning they will need to be cleaned/recoded.
-The rest of the fields are mostly classification variables and location information with some minor cleaning opportunities, but with 2,470,867 there is quite a bit of opportunity for frequency and classification analysis.
+My primary data source will come from [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data).
+The data is updated weekly by the Austin Police Department, and each record in the dataset represents an Incident Report, with the highest offense within an incident taking precedence in things like the description and categorization Each Incident can have other offense tied to it, however since each record is a unique incident then only the aforementioned Higehst Offense is the one represented (NOTE: At the time of this writing, 7/31/2024, the dataset has been taken down to be replaced with one that aligns more closely with the FBI National Incident Based Reporting System.
+The datasets are not one-to-one, so reproducibility would require more than a lift and shift, however some of the methods could still be employed).\
+The raw data is represented by several categorical, location, and time-based variables, many of which have missing values, or are formatted incorrectly for the data type, so they will need to be cleaned or recoded.
 After cleaning the data (done in a separate file found in this repo) we can take a look a more meaningful look.
 
-```{r Load Processed Data}
-d1 <- readRDS(here('data','processed-data','processedcrime.rds'))
+```{r Quick Summaries}
+skim(d1)
 ```
 
-```{r Quick Summaries}
-summary(d1)
-head(d1)
-```
+I preserved quite a few variables from the original dataset, but the primary ones of interest will be `Occurred.Date.Time`, `Zip.Code`, `Highest.Offense.Description`, `Category.Description`, and `Crime.Category`.
+The last few variables pertaining to the type of crime committed are really all rolled into the last one, `Crime.Category` for the purposes of this analysis.
+Crime Category is a derived field that categorizes crimes into several different categories for the sake of understanding what crime occurred but without getting flooded with minute details.
+The categorizations were done manually by myself, and can be seen in the processing file as well as a description of the method performed.
+The short version is the categories are based on the UCR Crime descriptions given by the FBI when available, and are bucketed similarly from the Highest Offense field via string detection.
+This brings the number of unique categories in the Highest Offense field from 436 to 32 in the derived field.
 
 ## Questions/Hypotheses to be addressed
+
 Ultimately, I would like to explore a few questions:
+
 1. Has crime truly dropped, or is it expected to rise again as the year progresses?
@@
 
-{{< pagebreak >}}
+## Schematic of workflow
+
+The intention of this analysis is to be exploratory; while I do have some questions I would like to chase down, I intend for them to be more of a compass than a map.
+With that said, the general strategy taken can be boiled down as follows:
+
+1. Check for data quality, understand data representation 
+2. Clean the data, address problems identified in previous step, some exploration 
+3. More "formal" exploration. Consider what variables might be most important, and how those variables can be represented in the analysis.
+4. Repeat any steps performed above as necessary.
+5. Resolve remaining or initial questions using statistical methods.
 
@@
 
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
+\
+To accomplish the steps described above, a number of techniques and methodologies were employed.
 
-## Schematic of workflow
+-   **Cleaning**: The data had a number of string variables for date fields, so converting these to proper date variables was necessary.
+    Several variables had missing or empty values.
+    Most of these records were removed, since they represented less than 1% of the data, and were often related to the seriousness of some crimes (like FBI descriptions and categorizations).
 
-```{r}
+-   **Imputing**: A few variables had the opportunity to impute the data logically.
+    For example, there were some records with a report date but no occurrence date.
+    By nature of the variable definitions, the occurrence date should always be less than or equal to the report date, so in instances when the occurrence time was missing but the report time was available, I imputed the occurrence date with the report date.
+    One could consider this a change in the definition of the variable, something like "the latest time the crime could have occurred," but for the purposes of this analysis that would suffice.
 
-```
+-   **Transformations/ Derived Variables**: The descriptions of the offenses were difficult to do much with simply because there were so many different types.
+    As a result, I ultimately decided to go through and bucket the different types of crimes by their description, by using a combination of other categorical variables when they were available and string detection to effectively create a lexicon that would categorize the variables.
 
-## Data aquisition
+-   **Exploratory Analysis**: The exploratory analysis done was very iterative, first focused the data quality, so a lot of filter, transforming, and summarizing the variables discussed above with basic summary and dplyer functions.
+    After cleaning visualizations like bar charts, choropleth charts, and line graphs were used to understand patterns in the data, largely done with ggplot and plotly.
+    After statistical analysis was performed, more visualizations in the form of line graphs with forecasts and error ranges.
 
-## Statistical analysis
-
-{{< pagebreak >}}
+-   **Statistical Analysis**: Primary statistical analysis performed was in the form of a hierarchical ARIMA forecast.
+    Exponential Smoothing was also briefly considered, as well as a non-hierarchical ARIMA method.
+    Primary methods to perform these forecasts were with the fable and forecast package.
+    The forecasts were produced using material from the the book Forecasting Principals and Practice (3rd Edition) [@hyndman2021].
 
@@
 
+This analysis focuses on three simple factors when it comes to the occurrence of crime: "when", "where", and "what".
+Quickly summarizing the results we see the following.\
+Starting with "When", we see that in recent years crime is generally decreasing and has been since around 2008.
+Year over year rates of homicide have been generally increasing in the years leading up to 2020.
+The claim made in the previously mentioned news article is that they are decreasing, and the time series plot arguably corroborates this[@fox2024].
+Still, the claim is worth investigating to see if the decreases are cyclical, seasonal, or just coincidence.
+Location variables are also provided in the data, so it is worth exploring the "where" that the crimes are occurring.
+Since zip code information was provided, and is a very simple location variable that people are familiar with, that seems like an easy place to start.
+We will see that crime is most common in a few zip codes in the center of the city, and that over time crime in those areas does appear to be decreasing, while the outer areas are relatively constant in the occurrence of crime.\
+Finally, the "what" is important simply because we can't celebrate decreases in crime if it's at the cost of more severe crime happening.
+We will isolate to homicide specifically, and observe the trends over time, then using data from 2003 through 2023 we will attempted to predict the month over month occurrence of homicide in 2024, to see how the data stacks up to what is actually happening.
+We ultimately fit a Hierarchical Seasonal ARIMA model, isolate to instances of Murder, and see that in the first few months of 2024, there may be a decrease in the rate of homicide compared to what was expected.
+
 ## Exploratory/Descriptive analysis
 
+\
+A time series plot is a very natural starting part for the exploration of this analysis.
+given that the data is at the incident level, it is worth the effort to aggregate the count of occurrences up to some interval of time.
+We have values of all the way down to reported time of occurrence, but given the imputations that needed to occur in the data cleaning step, as well as the questionable reliability of whether the reported occurrence time is accurate to begin with, daily seems like the smallest interval of time that could be worth aggregating over.
+Figure 1 below is the corresponding aggregated time series plot.
+While we can see a general trend, the plot is rather dense since there are so many points plotted along the x-axis.
+If we want to do anything like isolating down to specific crimes like homicide, we will needed to aggregate at less fine of a grain, otherwise we will have too many intervals with 0 values for the observations.
+
+![Fig 1](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/crime-by-day.png)  
+Figure 2 below is the corresponding time series plot aggregated over each month.
+We still see a similar rise and fall trend like we did in the daily plot, but it is more pronounced since there are fewer outliers like there were in the daily plot.
+Outside of the overall trend, we can also see some potential patterns like a seasonal trend, and a change in the overall trend around 2018 until around 2020 where it briefly increases again before returning to a general decrease.
+![Fig 2](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/crime-by-month.png)  
+Jumping ahead just a bit, after performing the simple lexicon based classification and properly identifying murder crimes, we can isolate to those crimes specifically for a time series plot.
+Figure 3 is exactly this, and we can see the trend is very different from what is seen in the overall time-series plot.
+For the most part, the rate of murder is relatively low up through 2015, and it is around 2016 that we see it spike and then slowly increase up until about 2022, where it then starts to slow back down.
+![Fig 3](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/murder-by-time.png)  
+These changes are interesting, primarily because we see the same shift before 2020 across overall crime and murder, but the number of murders is too small to be the primary driver in the overall crime trend.
+It's also interesting because of the way the increase in murders start happening before the overall increase, and continues after the overall has started to decrease again.
+These may be caused by correlated factors like population, economic, or political factors, that the rate of murder could be more sensitive to than other crimes are.
+Furthermore, because the rate of murder is so low (relative to other types of crime), it's likely that the variance in the trend is always going to be more sensitive to external factors.
+Once again the goal of this analysis is not intended to explain the why, but I think these observations could make for some good future studies into the potential causes or correlations.
+For now, the main question that these visuals bring me are the apparent trends; are they actually material, is the decrease seen the result of a change in the actual rate, or merely happenstance and we can expect the rate of murder to increase in future months?
+Before I attempt to explore this question, I want to see what variables might help me to explain it.
+The last chart considers the type of crime, which we will see in a moment, but it does not consider the location.
+For location, I have created a visualization based on Zip Code that I think can be revealing.
+
 ```{r}
-
+cplotly
 ```
 
-## Basic statistical analysis
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
+\
+The interactive plot above shows us the number of crimes that occurred each year within each zip code, with the number of crimes represented by the continuous color scale, and the years represented by the animation so that values will change as the animation plays.
+We can see the overall decrease in crime from the way the Zip Codes that are more red slowly start to shift to blue or white.
+The zip codes that are more blue to start with don't seem to change much as the animation plays.
+This suggests that the bulk of the crime is isolated to a handful of locations, and either factors that typically lead to crime are becoming less relevant in those areas, or the areas with heavier crime are seeing a greater focus in terms of prevention.
+This visualization is great, but with the analysis being more geared for a focus on Murder, it is unlikely to be of much help.
+There are probably Zip Codes where murder is more likely to occur, but the lower quantities would make it difficult to use that as a method to predict the number of murders that will happen in an interval of time small enough to have enough observations to make such a prediction in the first place.
+This chart does leave some questions that would be good for a future study, namely what has changed in those Zip Codes that are decreasing?
+Are there specific crimes that we are seeing fewer occurrences of, or is there better prevention in those areas?\
+The exploration has considered time and location as a variable, but the type of crime is still largely unexplored.
+The raw data, as stated before, has 434 different descriptions for the highest offense in each incident that was committed over the time period observed.
+The number of these descriptions makes it difficult to do anything with as a category, but coupled with other variables in the datset like the UCR Category may help us understand a better categorical representation.
+To accomplish this, crime that already had a UCR description for the observation was assigned this description as its category.
+A general description of each of these and several others can be found [here](https://ucr.fbi.gov/crime-in-the-u.s/2019/crime-in-the-u.s.-2019/topic-pages/offense-definitions)[@fbi2019].
+Observations in this dataset were given a UCR Category code when they were considered a more serious crime identified by the FBI (this was given in the data dictionary that came with the data.
+The data dictionary is no longer available since the dataset was removed, but can be found in the repo for this analysis on github in the assets folder, or from the processing file directly since I have it copied there).
+Because there were many observations that were not given a category in the variable field, a lexicon was built by observing each individual Offense descriptions to find a way to group each offense description into a smaller categorical variable.
+Make no mistake, this categorization is not official categorization of the crimes themselves, nor is it intended to be a one for one match to any other categorization used by either APD or the FBI.
+However, it can be interpreted as one person's layman's terms attempt at categorizing these offenses.
+Essentially, how might an average person who does not regularly encounter law enforcement or crime think of each of these offenses?
+With that in mind, after observing each of the Highest Offense description, key words and phrases were identified that could capture either unique or similar (when possible) descriptions of the crimes, then string detection and regular expressions were used to identify observations with those descripions and then group them into a category of other similar offenses.
+We can see exactly how each offense description was grouped in the table below.
 
 ```{r}
+RecatTable
 ```
 
-## Full analysis
+\
+The total count, yearly mean, and yearly standard deviation of occurrences of the offenses is provided to illustrate the disparity of the frequency of several of the offense descriptions.
+This is not entirely resolved by the New Category, but it does help for many of them.
+Again, the new categories are meant to generalize the descriptions.
+Page 21 of the table above contains the descriptions that I categorized as murder.
+A good example of how this method is imperfect is there, because of the presence of manslaughter.
+My initial thought process was to group instances of manslaughter into its own category, which is ultimately where I would have put Justified Homicide as well, however there were occurrences of manslaughter that were already categorized as murder by the UCR categorization, so bringing in those to the murder category would leave Justified Homicide in its own category, despite being a crime with a relatively low rate of occurrence.
+Ultimately this turned out to be the right call, because the FBI UCR website referenced earlier states that the "program classifies justifiable homicides separately,"[@fbi2019] but for other cases not explicitly stated on this or other sites I may have just gotten it wrong judging by my first instinct.
+All said, a more summarized version of this chart is below, complete with sparklines as well to give an idea of the rate over time of each category.
+
+```{=html}
+<iframe src="/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/html-widgets/new-cat-table.html" width="100%" height="800px" align="centered"></iframe>
+```
+\
+This table is intended to give a general idea of the change of each crime over time and the general quantity relative to others.
+We can again see that our specific crime of focus, Murder, is lower than many other types, but is far from the lowest and still has a quantifiable rate.
+The sparklines don't do the difference between crimes justice, because there are several that have a far greater rate of occurrence than others, but we can visualize this and the year over year change with the chart below.
 
 ```{r}
-
+catbarly
 ```
 
-{{< pagebreak >}}
+\
+Again, the rate of murder is generally so low that it is dwarfed by the occurrences of most other crimes (particularly theft).
+We can see the bar chart peak out towards the end in 2021 and beyond, which is exactly what we expect.
+Overall, the rate of murder is so low compared to others that we may not be able to give it the same treatment as others, but begs the question if the rate of crime as a whole could somehow be used to help us predict each individual type of crime.
+We will use this idea during the Forecasting steps described below.
 
-# Discussion
+## Time Series and Forecasting
 
-## Summary and Interpretation
+\
+To determine if the recent decreases seen in crime overall and murder are coincidence or not, we need to establish what our expectation is.
+There are many ways we could do this, but a forecast seems the best, given the time series nature of our data.
+A few different time series models were attempted, like ARIMA and Exponential Smoothing, but the confidence interval for these was too wide for any group to make any reasonable inference with.
+However, utilizing the Fable package, a top-down hierarchical Seasonal ARIMA model is fit to the data and gives us a much more reasonable fit, both for the aggregated level and for the individual crime level for Murder.
+To fit the model the specific model, automatic model selection in the fable package was utilized while reconciling with a top-down hierarchical method.
+The selected model for the aggregate level is an ARIMA(0,1,2)(2,1,1)~12~.
+We can see from the residual plots for All Crime that the fit is generally fair.
+Innovation residuals are mostly centered around 0, the ACF plot suggests no autocorrelation, and even the distribution of the residuals is pretty normal, though there is a high frequency at 0 and just below 0 that may be questionable.
+The Ljung-Box test was not found to be significant, meaning we can reject the null hypothesis that the residuals are distinguishable from white noise agreeing with our observations from the ACF plot.
+The most questionable aspect here is likely the heteroscedasticity of the residuals, which appears to decrease and then increase over time, but this is not a requirement for a good time series model.
+![Fig 4](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/crime-resid.png)
 
-## Strengths and Limitations
+```{=html}
+<iframe src="/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/aggregated-ljung.html" width="100%" height="250px" align="centered"></iframe>
+```
+\
+It is not quite the same story regarding the residuals for murder.
+The selected model for the frequency of murder is ARIMA(0,1,1)(0,0,1)~12~.
+The plot of the residuals over time is still mostly centered at zero but the variance is less steady than it was for all crime, increasing quite a bit in the last few years, which makes sense since that is when we started to see the rate of Murder increase.
+Two spikes in the ACF plot are found to be significant, but this is not terribly abormal since 24 spikes are plotted, and the Ljung-Box test results are also not found to be significant.
+The distribution of the residuals is the most questionable, it is fairly asymmetrical so it is unlikely to be normally distributed, but again this is less important than the requirements for uncorrelated and mean 0 residuals.
+![Fig 4](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/murd-resid.png)
 
-## Conclusions
+```{=html}
+<iframe src="/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/aggregated-ljung.html" width="100%" height="250px" align="centered"></iframe>
+```
+\
+The plots below show these forecasts, the aggregation all crime and for murder, as well as two other crimes picked for their similarity to murder, Aggregated Assault and Unlawful Restraint (kidnapping and similar crimes).
+Aggravated Assault was chosen due to its similarity in how the crime is performed, and Unlawful Restraint was chosen due to the number of occurrences being similar to Murder.
+The plot for Murder is repeated enlarged for readability.
 
-{{< pagebreak >}}
+![Fig 4](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/crime-forecast.png)  
+![Fig 5](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/murder-forecast.png)  
+From these forecasts, we can see that for overall crime the decrease observed earlier is expected to continue, however the confidence interval is quite wide so it could just as easily change course and begin to increase
+. For Murder, the decrease after the observed period is much less pronounced, but the confidence interval is much smaller
+.
+The difference in confidence interval is likely due to the smaller magnitude of crime counts compared to the aggregated model, and in this case the bounds are often not enough for more than one integer value between
+.
+That's somewhat problematic early on, because it means the forecast is truly only predicting one value since it is a count, but it matters very little since the last value is well outside of any of the confidence bounds
+. The table below shows the point estimates of the forecast versus the exact value, and we can see that the the estimates were not terribly bad except for March, where we see the large dip in the plot
+.
+It is worth noting that April and May were not plotted since the article that spawned the question at the heart of this analysis was only observing the count of murders through the first quarter of the year
+.
+
+```{=html}
+<iframe src="/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/actual-vs-forecast.html" width="100%" height="400px" align="centered"></iframe>
+```
+## Interpretation and Conlusions
+
+\
+The findings in the analysis are somewhat inconclusive, though they were really intended to be exploratory.
+There are some clear trends in crime rate over time, and forecasting methods do not look to be beyond the realm of possibility but should should certainly be taken with a grain of salt.
+There was a drop in the count of Murders in the first quarter, primarily due to a drop in the month of March.
+The forecast suggested it was expected to be higher, but we are talking about a difference of 2.7 actual vs forecast.
+The counts of murder are so low that there is bound to be error, and if not it would likely be due confidence intervals so wide they would be almost meaningless.
+with that in mind, the table above suggests that we have 5 fewer murders in the first 5 months of 2024 than expected, so cautious optimism may be warranted.\
+The model itself is certainly not perfect, and perhaps future analyses could explore ways to improve the predictability.
+Hierarchical time series models can have multiple levels in their hierarchy, so it's not impossible that other variables could have been combined with the category to improve results.
+Other crimes may also be more worthwhile in predicting; theft as an example dwarfed all other crimes and may very well be the biggest opportunity for the city of Austin.
+Models built on other higher frequency crimes can also likely make better use of location variables as well.
+More recent crime reporting data on the Austin data portal site have different methods of classification, and include individual charges with each incident, so it seems that they would be ripe for new variables to be implemented.
+Socio-economic variable could also potentially be used to see the impacts that they have on the rates of crime as whole or certain specific crimes.
+Comparisons with other cities would also be interesting, especially those with similarities to Austin to see how individual differences could lead to changes in the city's environment.
+These are all ideas I could see myself returning to this analyses for, and I hope this analyses can provide myself or anyone else some inspiration for a more detailed analysis.
 

### 2024-08-05 products/manuscript/Manuscript.qmd
staging after copy to portfolio

@@
 
-![Fig 1](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/crime-by-day.png)  
+```{r}
+#| label: fig-1
+#| fig-cap: "Crime Forecast Residual Plots"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures",'static-plots',"crime-by-day.png"))
+```
+
 Figure 2 below is the corresponding time series plot aggregated over each month.
@@
 Outside of the overall trend, we can also see some potential patterns like a seasonal trend, and a change in the overall trend around 2018 until around 2020 where it briefly increases again before returning to a general decrease.
-![Fig 2](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/crime-by-month.png)  
+
+```{r}
+#| label: fig-2
+#| fig-cap: "Crime Forecast Residual Plots"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures",'static-plots',"crime-by-month.png"))
+```
+
 Jumping ahead just a bit, after performing the simple lexicon based classification and properly identifying murder crimes, we can isolate to those crimes specifically for a time series plot.
@@
 For the most part, the rate of murder is relatively low up through 2015, and it is around 2016 that we see it spike and then slowly increase up until about 2022, where it then starts to slow back down.
-![Fig 3](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/murder-by-time.png)  
+
+```{r}
+#| label: fig-3
+#| fig-cap: "Monthly Murder Line Plot"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures",'static-plots',"murder-by-time.png"))
+```
+
 These changes are interesting, primarily because we see the same shift before 2020 across overall crime and murder, but the number of murders is too small to be the primary driver in the overall crime trend.
@@
 ```{=html}
-<iframe src="/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/html-widgets/new-cat-table.html" width="100%" height="800px" align="centered"></iframe>
+<iframe src="/HYLTIN-PII-project/results/figures/html-widgets/new-cat-table.html" width="100%" height="800px" align="centered"></iframe>
 ```
@@
 The most questionable aspect here is likely the heteroscedasticity of the residuals, which appears to decrease and then increase over time, but this is not a requirement for a good time series model.
-![Fig 4](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/crime-resid.png)
+
+```{r}
+#| label: fig-4
+#| fig-cap: "Crime Forecast Residual Plots"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures",'static-plots',"crime-resid.png"))
+```
 
 ```{=html}
-<iframe src="/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/aggregated-ljung.html" width="100%" height="250px" align="centered"></iframe>
+<iframe src="/HYLTIN-PII-project/results/figures/static-plots/aggregated-ljung.html" width="100%" height="250px" align="centered"></iframe>
 ```
@@
 The distribution of the residuals is the most questionable, it is fairly asymmetrical so it is unlikely to be normally distributed, but again this is less important than the requirements for uncorrelated and mean 0 residuals.
-![Fig 4](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/murd-resid.png)
+
+```{r}
+#| label: fig-5
+#| fig-cap: "Murder Forecast Residual Plots"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures",'static-plots',"murd-resid.png"))
+```
 
 ```{=html}
-<iframe src="/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/aggregated-ljung.html" width="100%" height="250px" align="centered"></iframe>
+<iframe src="/HYLTIN-PII-project/results/figures/static-plots/aggregated-ljung.html" width="100%" height="250px" align="centered"></iframe>
 ```
@@
 
-![Fig 4](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/crime-forecast.png)  
-![Fig 5](/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/murder-forecast.png)  
+```{r}
+#| label: fig-6
+#| fig-cap: "Monthly Crime Forecast"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures",'static-plots',"crime-forecast.png"))
+```
+
+```{r}
+#| label: fig-7
+#| fig-cap: "Murder Forecast"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures",'static-plots',"murder-forecast.png"))
+```
+
 From these forecasts, we can see that for overall crime the decrease observed earlier is expected to continue, however the confidence interval is quite wide so it could just as easily change course and begin to increase
@@
 ```{=html}
-<iframe src="/Users/Hylti/OneDrive/Documents/School/Practicum II/HYLTIN-PII-project/results/figures/static-plots/actual-vs-forecast.html" width="100%" height="400px" align="centered"></iframe>
+<iframe src="/HYLTIN-PII-project/results/figures/static-plots/actual-vs-forecast.html" width="100%" height="400px" align="centered"></iframe>
 ```

### 2024-08-05 code/analysis-code/analysisfile.qmd

@@
 ```{r}
+theme_set(theme_gray())
 aggresid <- group_fit %>% filter(is_aggregated(Crime.Category)) %>% select(td) %>% gg_tsresiduals()
@@
 ```{r}
+theme_set(theme_gray())
 murdresid <- group_fit %>% filter(Crime.Category == 'MURDER') %>% select(td) %>% gg_tsresiduals()
@@
 ```{r}
+theme_set(theme_minimal())
 agglb <- augment(filter(group_fit, is_aggregated(Crime.Category))) %>% 
@@
 ```{r}
+theme_set(theme_minimal())
 murdlb <- augment(filter(group_fit, Crime.Category == 'MURDER')) %>% 

### 2024-08-05 code/eda-code/eda.qmd

@@
   geom_line() +
-  labs(title = 'Austin Crime Over Time', x = 'Time of Occurence (by day)', y = 'Count')
+  labs(title = 'Austin Crime Over Time', x = 'Time of Occurence (by day)', y = 'Count') +
+  theme(panel.background = element_rect(fill = 'white',
+                                        color = 'white'),
+        plot.background = element_rect(fill = 'white',
+                                        color = 'white')
+        )
 
@@
   geom_line() +
-  labs(title = 'Austin Crime Over Time', x = 'Time of Occurence (by month)', y = 'Count')
+  labs(title = 'Austin Crime Over Time', x = 'Time of Occurence (by month)', y = 'Count') +
+  theme(panel.background = element_rect(fill = 'white',
+                                        color = 'white'),
+        plot.background = element_rect(fill = 'white',
+                                        color = 'white')
+        )
 
@@
   geom_line() +
-  labs(title = 'Austin Murder Over Time', x = 'Time of Occurence (by month)', y = 'Count')
+  labs(title = 'Austin Murder Over Time', x = 'Time of Occurence (by month)', y = 'Count') +
+  theme(panel.background = element_rect(fill = 'white',
+                                        color = 'white'),
+        plot.background = element_rect(fill = 'white',
+                                        color = 'white')
+        )
 

### 2024-08-06 products/manuscript/Manuscript.qmd
fixing project repo after submodule issues

@@
 ```{=html}
-<iframe src="/HYLTIN-PII-project/results/figures/html-widgets/new-cat-table.html" width="100%" height="800px" align="centered"></iframe>
+<iframe src="../../results/figures/html-widgets/new-cat-table.html" width="100%" height="800px" align="centered"></iframe>
 ```
@@
 ```{=html}
-<iframe src="/HYLTIN-PII-project/results/figures/static-plots/aggregated-ljung.html" width="100%" height="250px" align="centered"></iframe>
+<iframe src="../../results/figures/static-plots/aggregated-ljung.html" width="100%" height="250px" align="centered"></iframe>
 ```
@@
 ```{=html}
-<iframe src="/HYLTIN-PII-project/results/figures/static-plots/actual-vs-forecast.html" width="100%" height="400px" align="centered"></iframe>
+<iframe src="../../results/figures/static-plots/actual-vs-forecast.html" width="100%" height="400px" align="centered"></iframe>
 ```

### 2024-08-07 products/manuscript/Manuscript.qmd
last commit for final project

@@
 Aside from resources outlined in our course resources, I will also draw from this [book](https://otexts.com/fpp3/) [@hyndman2021].
-It's not one I have used before but it had been recommended to me previously and I would like to learn more about time series and forecasting in general.
+The final output of this project, the manuscript you are reading, is rendered in my github analytics [portfolio](https://whygit10.github.io/williamhyltin-P2-portfolio/).
+The original repo dedicated to just this project can be found [here](https://github.com/whygit10/HYLTIN-PII-project).
 
@@
 My primary data source will come from [here](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data).
-The data is updated weekly by the Austin Police Department, and each record in the dataset represents an Incident Report, with the highest offense within an incident taking precedence in things like the description and categorization Each Incident can have other offense tied to it, however since each record is a unique incident then only the aforementioned Higehst Offense is the one represented (NOTE: At the time of this writing, 7/31/2024, the dataset has been taken down to be replaced with one that aligns more closely with the FBI National Incident Based Reporting System.
+The data is updated weekly by the Austin Police Department, and each record in the dataset represents an Incident Report, with the highest offense within an incident taking precedence in things like the description and categorization Each Incident can have other offense tied to it, however since each record is a unique incident then only the aforementioned Higehst Offense is the one represented (NOTE: At the time of this writing, 7/31/2024, the dataset has been taken down to be [replaced](https://data.austintexas.gov/Public-Safety/NIBRS-Group-A-Offense-Crimes/i7fg-wrk5/about_data) with one that aligns more closely with the FBI National Incident Based Reporting System.
 The datasets are not one-to-one, so reproducibility would require more than a lift and shift, however some of the methods could still be employed).\
